---
title: "scFFPE Comprehensive DEG Analysis: ICI_AKI Signature Discovery"
author: "Kidney ICI_AKI Research Project"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
    code_folding: show
    theme: cosmo
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 12,
  fig.height = 8,
  dpi = 300
)
```

# ãƒ—ãƒ­ã‚¸ã‚§ã‚¯ãƒˆæ¦‚è¦

**ç›®çš„**: scFFPEãƒ‡ãƒ¼ã‚¿ã‹ã‚‰ICI-AKIç‰¹ç•°çš„ãªåˆ†å­ã‚·ã‚°ãƒãƒãƒ£ã‚’ä½“ç³»çš„ã«ç™ºè¦‹ã™ã‚‹

**å…¥åŠ›ãƒ‡ãƒ¼ã‚¿**: 
- File: `scffpe_additional_analysis_complete_doublet_removed.rds`
- Cells: Doubleté™¤åŽ»æ¸ˆã¿ã€è©³ç´°ã‚¢ãƒŽãƒ†ãƒ¼ã‚·ãƒ§ãƒ³å®Œäº†
- Samples: S0 (Rejection), S1/S3 (ICI-AKI), S2 (Drug-AKI)

**ã‚¢ãƒ—ãƒ­ãƒ¼ãƒ**: 
1. å…¨ç´°èƒžãƒ¬ãƒ™ãƒ«ã§ã®3ç¾¤æ¯”è¼ƒDEGè§£æž
2. ä¸»è¦ç´°èƒžã‚¿ã‚¤ãƒ—åˆ¥ã®DEGè§£æž
3. ICI-AKIç‰¹ç•°çš„éºä¼å­ã®åŒå®šï¼ˆä¸¡æ¯”è¼ƒã§æœ‰æ„ï¼‰
4. CosMx 1Kãƒ‘ãƒãƒ«ã¨ã®é‡è¤‡ç¢ºèª
5. Cell-state signatureã®å®šç¾©

**æœŸå¾…ã•ã‚Œã‚‹æˆæžœ**:
- ICIç‰¹ç•°çš„DEGãƒªã‚¹ãƒˆï¼ˆå…¨ç´°èƒž + å„ç´°èƒžã‚¿ã‚¤ãƒ—ï¼‰
- CosMxæŠ•å½±å¯èƒ½ãªã‚·ã‚°ãƒãƒãƒ£ï¼ˆ5-10éºä¼å­/stateï¼‰
- ç´°èƒžã‚¿ã‚¤ãƒ—åˆ¥ã®ç—…æ…‹ç‰¹ç•°çš„å¤‰åŒ–ã®è§£æ˜Ž

---

# 1. ç’°å¢ƒæº–å‚™ {.tabset .tabset-fade}

## ãƒ©ã‚¤ãƒ–ãƒ©ãƒªèª­ã¿è¾¼ã¿

```{r load_libraries}
cat("=== Loading Required Libraries ===\n")

suppressPackageStartupMessages({
  library(Seurat)
  library(SeuratObject)
  library(dplyr)
  library(tidyr)
  library(ggplot2)
  library(patchwork)
  library(viridis)
  library(RColorBrewer)
  library(pheatmap)
  library(readxl)
  library(ggrepel)
})

cat("âœ… All libraries loaded\n")
cat("Seurat version:", as.character(packageVersion("Seurat")), "\n")
cat("R version:", R.version.string, "\n\n")
```

## ãƒ‡ã‚£ãƒ¬ã‚¯ãƒˆãƒªè¨­å®š

```{r setup_directories}
cat("=== Setting up Directories ===\n")

# Project root
project_root <- "/Users/HiroyukiArai/Documents/å¤§å­¦é™¢/ç ”ç©¶çµæžœ/2024æ‘ä¸Šå…ˆç”ŸCosMX/202406 single cell FFPE data/scFFPE_kidney_analysis"

# Data directory
data_dir <- file.path(project_root, "results")

# Output directory
output_dir <- file.path(project_root, "results/DEG_analysis")
dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)

# CosMx panel file
cosmx_panel_file <- "/Users/HiroyukiArai/Documents/å¤§å­¦é™¢/ç ”ç©¶çµæžœ/2024æ‘ä¸Šå…ˆç”ŸCosMX/CosMxâ„¢ Human Universal Cell Characterization Panel Gene Target List.xlsx"

cat("Data directory:", data_dir, "\n")
cat("Output directory:", output_dir, "\n")
cat("CosMx panel file:", cosmx_panel_file, "\n\n")
```

## ãƒ‡ãƒ¼ã‚¿èª­ã¿è¾¼ã¿

```{r load_data}
cat("=== Loading scFFPE Data ===\n")

# Load scFFPE object
scffpe_file <- file.path(data_dir, "scffpe_additional_analysis_complete_doublet_removed.rds")

if(!file.exists(scffpe_file)) {
  stop("âŒ Data file not found: ", scffpe_file)
}

cat("Loading data...\n")
scffpe <- readRDS(scffpe_file)

cat("âœ… Data loaded successfully\n")
cat("Total cells:", ncol(scffpe), "\n")
cat("Total genes:", nrow(scffpe), "\n")
cat("Samples:", paste(unique(scffpe$orig.ident), collapse = ", "), "\n")
cat("Assays:", names(scffpe@assays), "\n\n")

# Check metadata
cat("Available metadata columns:\n")
print(colnames(scffpe@meta.data))
cat("\n")
```

## Groupå®šç¾©

```{r define_groups}
cat("=== Defining Sample Groups ===\n")

# Check current sample info
cat("Current sample distribution:\n")
print(table(scffpe$orig.ident))

# Create group column based on sample
# First check what column contains sample info
if("sample" %in% colnames(scffpe@meta.data)) {
  sample_col <- "sample"
} else if("orig.ident" %in% colnames(scffpe@meta.data)) {
  sample_col <- "orig.ident"
} else {
  stop("Cannot find sample information column")
}

# Check actual sample names
cat("\nActual sample names in data:\n")
print(unique(scffpe@meta.data[[sample_col]]))

# Create group based on actual sample names (with _scFFPE suffix)
scffpe$group <- case_when(
  scffpe@meta.data[[sample_col]] == "S0_scFFPE" ~ "Rejection",
  scffpe@meta.data[[sample_col]] %in% c("S1_scFFPE", "S3_scFFPE") ~ "ICI_AKI",
  scffpe@meta.data[[sample_col]] == "S2_scFFPE" ~ "Drug_AKI",
  TRUE ~ "Unknown"
)

# Verify grouping
cat("\nâœ… Group assignment complete:\n")
print(table(scffpe$group))

cat("\nFinal group distribution:\n")
print(table(scffpe$group))
cat("\n")
```

## CosMxãƒ‘ãƒãƒ«èª­ã¿è¾¼ã¿

```{r load_cosmx_panel}
cat("=== Loading CosMx Gene Panel ===\n")

if(!file.exists(cosmx_panel_file)) {
  cat("âš ï¸  CosMx panel file not found, will skip overlap analysis\n")
  cosmx_genes <- NULL
} else {
  # Read gene list from Annotations sheet
  cosmx_panel_df <- read_excel(cosmx_panel_file, sheet = "Annotations", skip = 1)
  
  # Extract gene names
  cosmx_genes <- cosmx_panel_df[[1]]
  cosmx_genes <- cosmx_genes[!is.na(cosmx_genes)]
  cosmx_genes <- cosmx_genes[cosmx_genes != "Gene"]
  
  cat("âœ… CosMx panel genes loaded\n")
  cat("Total panel genes:", length(cosmx_genes), "\n")
  
  # Check overlap with scFFPE data
  genes_in_scffpe <- intersect(cosmx_genes, rownames(scffpe))
  cat("Genes available in scFFPE:", length(genes_in_scffpe), 
      sprintf("(%.1f%%)\n\n", length(genes_in_scffpe)/length(cosmx_genes)*100))
}
```

---

# 2. å…¨ç´°èƒžãƒ¬ãƒ™ãƒ«DEGè§£æž {.tabset .tabset-fade}

## ICI-AKI vs Rejection

```{r deg_ici_vs_rejection}
cat("=== DEG Analysis: ICI-AKI vs Rejection (All Cells) ===\n")

# Set identity
Idents(scffpe) <- "group"

# FindMarkers
cat("Running FindMarkers (this may take several minutes)...\n")
deg_ici_vs_rej <- FindMarkers(
  scffpe,
  ident.1 = "ICI_AKI",
  ident.2 = "Rejection",
  logfc.threshold = 0.25,
  min.pct = 0.1,
  test.use = "wilcox"
)

# Add gene names as column
deg_ici_vs_rej$gene <- rownames(deg_ici_vs_rej)

# Filter significant genes
deg_ici_vs_rej_sig <- deg_ici_vs_rej %>%
  filter(p_val_adj < 0.05) %>%
  mutate(
    direction = ifelse(avg_log2FC > 0, "ICI_high", "Rejection_high"),
    abs_log2FC = abs(avg_log2FC)
  ) %>%
  arrange(desc(abs_log2FC))

cat("âœ… DEG analysis complete\n")
cat("Total genes tested:", nrow(deg_ici_vs_rej), "\n")
cat("Significant genes (padj < 0.05):", nrow(deg_ici_vs_rej_sig), "\n")
cat("  - ICI_AKI high:", sum(deg_ici_vs_rej_sig$direction == "ICI_high"), "\n")
cat("  - Rejection high:", sum(deg_ici_vs_rej_sig$direction == "Rejection_high"), "\n\n")

# Save results
write.csv(deg_ici_vs_rej_sig, 
          file.path(output_dir, "DEG_ICI_vs_Rejection_all_cells.csv"),
          row.names = FALSE)

# Show top genes
cat("Top 20 ICI_high genes:\n")
print(head(deg_ici_vs_rej_sig %>% filter(direction == "ICI_high"), 20))

cat("\nTop 20 Rejection_high genes:\n")
print(head(deg_ici_vs_rej_sig %>% filter(direction == "Rejection_high"), 20))
```

## ICI-AKI vs Drug-AKI

```{r deg_ici_vs_drug}
cat("=== DEG Analysis: ICI-AKI vs Drug-AKI (All Cells) ===\n")

# FindMarkers
cat("Running FindMarkers (this may take several minutes)...\n")
deg_ici_vs_drug <- FindMarkers(
  scffpe,
  ident.1 = "ICI_AKI",
  ident.2 = "Drug_AKI",
  logfc.threshold = 0.25,
  min.pct = 0.1,
  test.use = "wilcox"
)

# Add gene names
deg_ici_vs_drug$gene <- rownames(deg_ici_vs_drug)

# Filter significant genes
deg_ici_vs_drug_sig <- deg_ici_vs_drug %>%
  filter(p_val_adj < 0.05) %>%
  mutate(
    direction = ifelse(avg_log2FC > 0, "ICI_high", "Drug_high"),
    abs_log2FC = abs(avg_log2FC)
  ) %>%
  arrange(desc(abs_log2FC))

cat("âœ… DEG analysis complete\n")
cat("Total genes tested:", nrow(deg_ici_vs_drug), "\n")
cat("Significant genes (padj < 0.05):", nrow(deg_ici_vs_drug_sig), "\n")
cat("  - ICI_AKI high:", sum(deg_ici_vs_drug_sig$direction == "ICI_high"), "\n")
cat("  - Drug-AKI high:", sum(deg_ici_vs_drug_sig$direction == "Drug_high"), "\n\n")

# Save results
write.csv(deg_ici_vs_drug_sig, 
          file.path(output_dir, "DEG_ICI_vs_Drug_all_cells.csv"),
          row.names = FALSE)

# Show top genes
cat("Top 20 ICI_high genes:\n")
print(head(deg_ici_vs_drug_sig %>% filter(direction == "ICI_high"), 20))

cat("\nTop 20 Drug_high genes:\n")
print(head(deg_ici_vs_drug_sig %>% filter(direction == "Drug_high"), 20))
```

## ICIç‰¹ç•°çš„éºä¼å­

```{r ici_specific_genes}
cat("=== Identifying ICI-Specific Genes (Both Comparisons) ===\n")

# Extract ICI_high genes from both comparisons
ici_high_vs_rej <- deg_ici_vs_rej_sig %>% 
  filter(direction == "ICI_high") %>%
  select(gene, log2FC_vs_Rej = avg_log2FC, padj_vs_Rej = p_val_adj)

ici_high_vs_drug <- deg_ici_vs_drug_sig %>% 
  filter(direction == "ICI_high") %>%
  select(gene, log2FC_vs_Drug = avg_log2FC, padj_vs_Drug = p_val_adj)

# Inner join (both comparisons)
ici_specific <- inner_join(ici_high_vs_rej, ici_high_vs_drug, by = "gene") %>%
  mutate(
    mean_log2FC = (log2FC_vs_Rej + log2FC_vs_Drug) / 2,
    max_padj = pmax(padj_vs_Rej, padj_vs_Drug)
  ) %>%
  arrange(desc(mean_log2FC))

cat("âœ… ICI-specific gene identification complete\n")
cat("Genes high in ICI_AKI vs Rejection:", nrow(ici_high_vs_rej), "\n")
cat("Genes high in ICI_AKI vs Drug-AKI:", nrow(ici_high_vs_drug), "\n")
cat("ICI-specific genes (both comparisons):", nrow(ici_specific), "\n\n")

# Save results
write.csv(ici_specific, 
          file.path(output_dir, "ICI_specific_genes_all_cells.csv"),
          row.names = FALSE)

# Show top ICI-specific genes
cat("Top 50 ICI-specific genes:\n")
print(head(ici_specific, 50))
```

## Volcano Plots

```{r volcano_plots, fig.width=14, fig.height=6}
cat("=== Creating Volcano Plots ===\n")

# Prepare data for volcano plots
deg_ici_vs_rej_volcano <- deg_ici_vs_rej %>%
  mutate(
    significance = case_when(
      p_val_adj < 0.05 & avg_log2FC > 0.5 ~ "ICI_high",
      p_val_adj < 0.05 & avg_log2FC < -0.5 ~ "Rejection_high",
      TRUE ~ "NS"
    )
  )

deg_ici_vs_drug_volcano <- deg_ici_vs_drug %>%
  mutate(
    significance = case_when(
      p_val_adj < 0.05 & avg_log2FC > 0.5 ~ "ICI_high",
      p_val_adj < 0.05 & avg_log2FC < -0.5 ~ "Drug_high",
      TRUE ~ "NS"
    )
  )

# Top genes to label
top_genes_rej <- deg_ici_vs_rej_volcano %>%
  filter(significance != "NS") %>%
  arrange(desc(abs(avg_log2FC))) %>%
  head(10) %>%
  pull(gene)

top_genes_drug <- deg_ici_vs_drug_volcano %>%
  filter(significance != "NS") %>%
  arrange(desc(abs(avg_log2FC))) %>%
  head(10) %>%
  pull(gene)

# Plot 1: ICI vs Rejection
p1 <- ggplot(deg_ici_vs_rej_volcano, 
             aes(x = avg_log2FC, y = -log10(p_val_adj), color = significance)) +
  geom_point(alpha = 0.6, size = 1.5) +
  geom_text_repel(
    data = deg_ici_vs_rej_volcano %>% filter(gene %in% top_genes_rej),
    aes(label = gene),
    size = 3,
    max.overlaps = 20
  ) +
  scale_color_manual(values = c("ICI_high" = "#E64B35", 
                                  "Rejection_high" = "#4DBBD5", 
                                  "NS" = "grey70")) +
  geom_vline(xintercept = c(-0.5, 0.5), linetype = "dashed", color = "grey40") +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed", color = "grey40") +
  labs(title = "DEG: ICI_AKI vs Rejection (All Cells)",
       x = "log2 Fold Change",
       y = "-log10(adjusted p-value)",
       color = "Significance") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom")

print(p1)

ggsave(file.path(output_dir, "Volcano_ICI_vs_Rejection_all_cells.png"),
       p1, width = 10, height = 8, dpi = 300)

# Plot 2: ICI vs Drug
p2 <- ggplot(deg_ici_vs_drug_volcano, 
             aes(x = avg_log2FC, y = -log10(p_val_adj), color = significance)) +
  geom_point(alpha = 0.6, size = 1.5) +
  geom_text_repel(
    data = deg_ici_vs_drug_volcano %>% filter(gene %in% top_genes_drug),
    aes(label = gene),
    size = 3,
    max.overlaps = 20
  ) +
  scale_color_manual(values = c("ICI_high" = "#E64B35", 
                                  "Drug_high" = "#00A087", 
                                  "NS" = "grey70")) +
  geom_vline(xintercept = c(-0.5, 0.5), linetype = "dashed", color = "grey40") +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed", color = "grey40") +
  labs(title = "DEG: ICI_AKI vs Drug-AKI (All Cells)",
       x = "log2 Fold Change",
       y = "-log10(adjusted p-value)",
       color = "Significance") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom")

print(p2)

ggsave(file.path(output_dir, "Volcano_ICI_vs_Drug_all_cells.png"),
       p2, width = 10, height = 8, dpi = 300)

cat("âœ… Volcano plots saved\n\n")
```

---

# 3. ç´°èƒžã‚¿ã‚¤ãƒ—åˆ¥DEGè§£æž {.tabset .tabset-fade}

## ç´°èƒžã‚¿ã‚¤ãƒ—ç¢ºèª

```{r check_celltypes}
cat("=== Checking Available Cell Types ===\n")

# Check if major_celltype exists
if("major_celltype" %in% colnames(scffpe@meta.data)) {
  cat("âœ… major_celltype column found\n")
  celltype_col <- "major_celltype"
} else if("detailed_celltype" %in% colnames(scffpe@meta.data)) {
  cat("âš ï¸  Using detailed_celltype instead\n")
  celltype_col <- "detailed_celltype"
} else {
  stop("âŒ No cell type annotation found in metadata")
}

# Cell type distribution
cat("\nCell type distribution:\n")
celltype_table <- table(scffpe@meta.data[[celltype_col]])
print(sort(celltype_table, decreasing = TRUE))

# Cell type by group
cat("\nCell type distribution by group:\n")
celltype_group_table <- table(scffpe@meta.data[[celltype_col]], scffpe$group)
print(celltype_group_table)

# Select cell types with sufficient cells (>50 per group minimum)
sufficient_celltypes <- names(celltype_table)[celltype_table >= 100]

cat("\nâœ… Cell types with sufficient cells (â‰¥100):", length(sufficient_celltypes), "\n")
print(sufficient_celltypes)
cat("\n")
```

## ç´°èƒžã‚¿ã‚¤ãƒ—åˆ¥DEGå®Ÿè¡Œ

```{r celltype_deg_analysis}
cat("=== Cell Type-Specific DEG Analysis ===\n")

# Initialize results storage
celltype_deg_results <- list()

# Loop through each cell type
for(ct in sufficient_celltypes) {
  
  cat("\n", strrep("=", 70), "\n")
  cat("Analyzing:", ct, "\n")
  cat(strrep("=", 70), "\n")
  
# Subset cells (correct method for Seurat v5)
cells_subset <- subset(scffpe, subset = !!sym(celltype_col) == ct)
  
  # Check cell counts per group
  group_counts <- table(cells_subset$group)
  cat("Cell counts per group:\n")
  print(group_counts)
  
  # Skip if insufficient cells in any group
  if(any(group_counts < 10)) {
    cat("âš ï¸  Skipped: insufficient cells in one or more groups\n")
    next
  }
  
  # Set identity
  Idents(cells_subset) <- "group"
  
  # === ICI vs Rejection ===
  cat("\nRunning: ICI_AKI vs Rejection...\n")
  
  tryCatch({
    deg_ct_rej <- FindMarkers(
      cells_subset,
      ident.1 = "ICI_AKI",
      ident.2 = "Rejection",
      logfc.threshold = 0.25,
      min.pct = 0.1,
      test.use = "wilcox"
    )
    
    deg_ct_rej$gene <- rownames(deg_ct_rej)
    deg_ct_rej$celltype <- ct
    deg_ct_rej$comparison <- "ICI_vs_Rejection"
    
    deg_ct_rej_sig <- deg_ct_rej %>%
      filter(p_val_adj < 0.05) %>%
      mutate(direction = ifelse(avg_log2FC > 0, "ICI_high", "Rejection_high"))
    
    cat("  Significant DEGs:", nrow(deg_ct_rej_sig), "\n")
    
  }, error = function(e) {
    cat("  âŒ Error:", e$message, "\n")
    deg_ct_rej_sig <- NULL
  })
  
  # === ICI vs Drug ===
  cat("Running: ICI_AKI vs Drug-AKI...\n")
  
  tryCatch({
    deg_ct_drug <- FindMarkers(
      cells_subset,
      ident.1 = "ICI_AKI",
      ident.2 = "Drug_AKI",
      logfc.threshold = 0.25,
      min.pct = 0.1,
      test.use = "wilcox"
    )
    
    deg_ct_drug$gene <- rownames(deg_ct_drug)
    deg_ct_drug$celltype <- ct
    deg_ct_drug$comparison <- "ICI_vs_Drug"
    
    deg_ct_drug_sig <- deg_ct_drug %>%
      filter(p_val_adj < 0.05) %>%
      mutate(direction = ifelse(avg_log2FC > 0, "ICI_high", "Drug_high"))
    
    cat("  Significant DEGs:", nrow(deg_ct_drug_sig), "\n")
    
  }, error = function(e) {
    cat("  âŒ Error:", e$message, "\n")
    deg_ct_drug_sig <- NULL
  })
  
  # Store results
  if(!is.null(deg_ct_rej_sig) || !is.null(deg_ct_drug_sig)) {
    celltype_deg_results[[ct]] <- list(
      vs_Rejection = deg_ct_rej_sig,
      vs_Drug = deg_ct_drug_sig
    )
  }
}

cat("\n", strrep("=", 70), "\n")
cat("âœ… Cell type-specific DEG analysis complete\n")
cat("Analyzed cell types:", length(celltype_deg_results), "\n")
cat(strrep("=", 70), "\n\n")

# Save results
saveRDS(celltype_deg_results, 
        file.path(output_dir, "celltype_specific_DEG_results.rds"))
```

## çµ±åˆçµæžœã¾ã¨ã‚

```{r celltype_deg_summary}
cat("=== Summarizing Cell Type-Specific DEGs ===\n")

if(length(celltype_deg_results) == 0) {
  cat("âš ï¸  No cell type-specific results available\n")
} else {
  
  # Combine all results
  all_celltype_deg <- bind_rows(
    lapply(names(celltype_deg_results), function(ct) {
      bind_rows(
        celltype_deg_results[[ct]]$vs_Rejection,
        celltype_deg_results[[ct]]$vs_Drug
      )
    })
  )
  
  cat("âœ… Combined results\n")
  cat("Total cell type Ã— gene Ã— comparison combinations:", nrow(all_celltype_deg), "\n\n")
  
  # Save combined results
  write.csv(all_celltype_deg, 
            file.path(output_dir, "All_celltype_specific_DEGs.csv"),
            row.names = FALSE)
  
  # Summary by cell type
  cat("Summary by cell type:\n")
  summary_by_ct <- all_celltype_deg %>%
    group_by(celltype, comparison) %>%
    summarize(
      n_genes = n(),
      n_ici_high = sum(direction %in% c("ICI_high")),
      n_other_high = sum(!direction %in% c("ICI_high")),
      .groups = "drop"
    )
  print(summary_by_ct)
  
  # Top DEGs per cell type
  cat("\n\nTop 5 DEGs per cell type (ICI_high):\n")
  top_degs_per_ct <- all_celltype_deg %>%
    filter(direction == "ICI_high") %>%
    group_by(celltype, comparison) %>%
    arrange(desc(abs(avg_log2FC))) %>%
    slice_head(n = 5) %>%
    select(celltype, comparison, gene, avg_log2FC, p_val_adj)
  
  print(as.data.frame(top_degs_per_ct))
}

cat("\n")
```

## DEG Heatmap

```{r celltype_deg_heatmap, fig.width=12, fig.height=10}
if(length(celltype_deg_results) > 0 && nrow(all_celltype_deg) > 0) {
  
  cat("=== Creating Cell Type DEG Heatmap ===\n")
  
  # Select top genes per cell type
  top_genes_per_ct <- all_celltype_deg %>%
    filter(direction == "ICI_high") %>%
    group_by(celltype) %>%
    arrange(desc(abs(avg_log2FC))) %>%
    slice_head(n = 10) %>%
    pull(gene) %>%
    unique()
  
  cat("Selected genes for heatmap:", length(top_genes_per_ct), "\n")
  
  if(length(top_genes_per_ct) > 0) {
    
    # Create matrix for heatmap
heatmap_data <- all_celltype_deg %>%
  filter(gene %in% top_genes_per_ct, comparison == "ICI_vs_Rejection") %>%
  select(celltype, gene, avg_log2FC) %>%
  pivot_wider(names_from = celltype, values_from = avg_log2FC, values_fill = 0)

    # Convert to data frame with rownames
    heatmap_matrix <- as.data.frame(heatmap_data)
    rownames(heatmap_matrix) <- heatmap_matrix$gene
    heatmap_matrix$gene <- NULL
    
    # Plot heatmap
    pheatmap(
      as.matrix(heatmap_matrix),
      cluster_rows = TRUE,
      cluster_cols = TRUE,
      color = colorRampPalette(c("blue", "white", "red"))(100),
      breaks = seq(-2, 2, length.out = 101),
      main = "Top DEGs per Cell Type (ICI_AKI vs Rejection)",
      fontsize = 8,
      fontsize_row = 7,
      fontsize_col = 9,
      filename = file.path(output_dir, "Celltype_DEG_heatmap.pdf"),
      width = 12,
      height = 10
    )
    
    cat("âœ… Heatmap saved\n")
  }
}

cat("\n")
```

---

# 4. CosMx 1Kãƒ‘ãƒãƒ«é‡è¤‡ç¢ºèª {.tabset .tabset-fade}

## å…¨ç´°èƒžDEGé‡è¤‡

```{r cosmx_overlap_all_cells}
cat("=== CosMx Panel Overlap: All Cells ===\n")

if(is.null(cosmx_genes)) {
  cat("âš ï¸  CosMx panel not loaded, skipping overlap analysis\n")
} else {
  
  # === ICI-specific genes overlap ===
  cat("ICI-specific genes (both comparisons):\n")
  
  ici_specific_cosmx <- ici_specific %>%
    mutate(in_CosMx = gene %in% cosmx_genes) %>%
    arrange(desc(mean_log2FC))
  
  cat("Total ICI-specific genes:", nrow(ici_specific_cosmx), "\n")
  cat("Available in CosMx panel:", sum(ici_specific_cosmx$in_CosMx), 
      sprintf("(%.1f%%)\n\n", sum(ici_specific_cosmx$in_CosMx)/nrow(ici_specific_cosmx)*100))
  
  # Save overlap results
  write.csv(ici_specific_cosmx, 
            file.path(output_dir, "ICI_specific_genes_CosMx_overlap.csv"),
            row.names = FALSE)
  
  # Show top genes available in CosMx
  cat("Top 30 ICI-specific genes available in CosMx:\n")
  top_cosmx <- ici_specific_cosmx %>%
    filter(in_CosMx == TRUE) %>%
    head(30)
  print(top_cosmx)
  
  # === Summary by FC ranges ===
  cat("\n\nCosMx overlap by fold change range:\n")
  fc_ranges <- ici_specific_cosmx %>%
    mutate(
      FC_range = cut(mean_log2FC, 
                     breaks = c(0, 0.5, 1.0, 1.5, 2.0, Inf),
                     labels = c("0-0.5", "0.5-1.0", "1.0-1.5", "1.5-2.0", ">2.0"))
    ) %>%
    group_by(FC_range) %>%
    summarize(
      total = n(),
      in_CosMx = sum(in_CosMx),
      pct_in_CosMx = round(sum(in_CosMx)/n()*100, 1)
    )
  print(fc_ranges)
}

cat("\n")
```

## ç´°èƒžã‚¿ã‚¤ãƒ—åˆ¥é‡è¤‡

```{r cosmx_overlap_celltype}
cat("=== CosMx Panel Overlap: Cell Type-Specific ===\n")

if(is.null(cosmx_genes) || length(celltype_deg_results) == 0) {
  cat("âš ï¸  CosMx panel or cell type results not available\n")
} else {
  
  # Initialize overlap storage
  celltype_cosmx_overlap <- list()
  
  # Loop through cell types
  for(ct in names(celltype_deg_results)) {
    
    cat("\n", ct, ":\n")
    
    # vs Rejection
    if(!is.null(celltype_deg_results[[ct]]$vs_Rejection)) {
      rej_genes <- celltype_deg_results[[ct]]$vs_Rejection %>%
        filter(direction == "ICI_high") %>%
        pull(gene)
      rej_cosmx <- intersect(rej_genes, cosmx_genes)
      
      cat("  vs Rejection: ", length(rej_cosmx), "/", length(rej_genes), 
          sprintf(" (%.1f%%) in CosMx\n", length(rej_cosmx)/length(rej_genes)*100))
      
      celltype_cosmx_overlap[[ct]]$vs_Rejection_CosMx <- rej_cosmx
    }
    
    # vs Drug
    if(!is.null(celltype_deg_results[[ct]]$vs_Drug)) {
      drug_genes <- celltype_deg_results[[ct]]$vs_Drug %>%
        filter(direction == "ICI_high") %>%
        pull(gene)
      drug_cosmx <- intersect(drug_genes, cosmx_genes)
      
      cat("  vs Drug-AKI: ", length(drug_cosmx), "/", length(drug_genes), 
          sprintf(" (%.1f%%) in CosMx\n", length(drug_cosmx)/length(drug_genes)*100))
      
      celltype_cosmx_overlap[[ct]]$vs_Drug_CosMx <- drug_cosmx
    }
  }
  
  # Save overlap results
  saveRDS(celltype_cosmx_overlap, 
          file.path(output_dir, "celltype_DEG_CosMx_overlap.rds"))
  
  # Summary table
  cat("\n\n=== Summary Table ===\n")
  
  overlap_summary <- data.frame(
    Cell_Type = names(celltype_cosmx_overlap),
    vs_Rejection_CosMx = sapply(celltype_cosmx_overlap, function(x) length(x$vs_Rejection_CosMx)),
    vs_Drug_CosMx = sapply(celltype_cosmx_overlap, function(x) length(x$vs_Drug_CosMx))
  )
  
  print(overlap_summary)
  
  write.csv(overlap_summary, 
            file.path(output_dir, "celltype_CosMx_overlap_summary.csv"),
            row.names = FALSE)
  
  cat("\nâœ… Cell type-specific CosMx overlap analysis complete\n")
}

cat("\n")
```

## Signatureå€™è£œä½œæˆ

```{r signature_candidates}
# ========================================
# Signature Candidates Creation (IMPROVED)
# ========================================

cat("=== Creating Signature Candidates for CosMx Projection ===\n")

if(is.null(cosmx_genes) || length(celltype_deg_results) == 0) {
  cat("âš ï¸  Required data not available\n")
} else {
  
  # ========================================
  # Strategy 1: Cell Type-Specific Signatures
  # ========================================
  
  cat("\n--- Strategy 1: Cell Type-Specific Signatures ---\n")
  
  signature_definitions <- list()
  
  # Function: Extract top CosMx-available genes per cell type
  extract_celltype_signature <- function(celltype_name, comparison, n_genes = 10) {
    if(!celltype_name %in% names(celltype_deg_results)) return(character(0))
    
    if(comparison == "vs_Rejection") {
      degs <- celltype_deg_results[[celltype_name]]$vs_Rejection
    } else {
      degs <- celltype_deg_results[[celltype_name]]$vs_Drug
    }
    
    if(is.null(degs) || nrow(degs) == 0) return(character(0))
    
    # Filter ICI-high + CosMx available
    sig_genes <- degs %>%
      filter(direction == "ICI_high") %>%
      filter(gene %in% cosmx_genes) %>%
      arrange(desc(abs(avg_log2FC))) %>%
      head(n_genes) %>%
      pull(gene)
    
    return(sig_genes)
  }
  
  # Tubular Injury Signature
  signature_definitions$Tubular_Injury <- extract_celltype_signature(
    "Tubular_cells", "vs_Rejection", 10
  )
  
  # Endothelial Activation Signature
  signature_definitions$Endothelial_Activation <- extract_celltype_signature(
    "Endothelial_cells", "vs_Rejection", 10
  )
  
  # Myeloid Activation Signature
  signature_definitions$Myeloid_Activation <- extract_celltype_signature(
    "Myeloid_cells", "vs_Rejection", 10
  )
  
  # T cell Activation Signature
  signature_definitions$T_cell_Activation <- extract_celltype_signature(
    "T_NK_cells", "vs_Rejection", 10
  )
  
  # Stromal Activation Signature
  signature_definitions$Stromal_Activation <- extract_celltype_signature(
    "Stromal_cells", "vs_Rejection", 10
  )
  
  # B/Plasma cell Signature
  signature_definitions$B_Plasma_Activation <- extract_celltype_signature(
    "B_Plasma_cells", "vs_Rejection", 8
  )
  
  # ========================================
  # Strategy 2: Pathway-Based Signatures (Manual Curation)
  # ========================================
  
  cat("\n--- Strategy 2: Pathway-Based Signatures (Manual) ---\n")
  
  # IFN-response (manual curation from known genes)
  ifn_candidates <- c("IFIT1", "IFIT2", "IFIT3", "ISG15", "ISG20", 
                      "MX1", "MX2", "OAS1", "OAS2", "OAS3",
                      "STAT1", "STAT2", "IRF1", "IRF7", "IRF9",
                      "IFI6", "IFI44", "IFI44L", "IFITM1", "IFITM3",
                      "GBP1", "GBP2", "GBP4", "GBP5", "CXCL10", "CXCL9")
  
  ifn_available <- intersect(ifn_candidates, 
                              ici_specific_cosmx %>% 
                                filter(in_CosMx == TRUE) %>% 
                                pull(gene))
  
  if(length(ifn_available) > 0) {
    signature_definitions$IFN_Response <- ifn_available
  }
  
  # Inflammatory/Chemokine Signature
  chemokine_candidates <- c("CCL2", "CCL3", "CCL4", "CCL5", "CCL7", "CCL8",
                            "CCL20", "CXCL1", "CXCL2", "CXCL3", "CXCL8",
                            "CXCL9", "CXCL10", "CXCL11", "CXCL12", "CXCL13", "CXCL14",
                            "IL1B", "IL6", "IL8", "IL17RB", "TNF")
  
  chemokine_available <- intersect(chemokine_candidates,
                                   ici_specific_cosmx %>% 
                                     filter(in_CosMx == TRUE) %>% 
                                     pull(gene))
  
  if(length(chemokine_available) >= 3) {
    signature_definitions$Inflammatory_Chemokine <- chemokine_available
  }
  
  # Exhaustion Signature (T cell specific)
  exhaustion_candidates <- c("PDCD1", "CTLA4", "LAG3", "TIGIT", "HAVCR2",
                             "TOX", "TOX2", "ENTPD1", "LAYN", "CD38",
                             "CXCL13", "IFNG", "GZMB", "PRF1")
  
  exhaustion_available <- intersect(exhaustion_candidates,
                                    cosmx_genes)  # Check in full CosMx panel
  
  # Extract from T cell DEGs specifically
  if("T_NK_cells" %in% names(celltype_deg_results)) {
    t_degs <- celltype_deg_results$T_NK_cells$vs_Rejection
    if(!is.null(t_degs)) {
      exhaustion_t <- t_degs %>%
        filter(gene %in% exhaustion_available) %>%
        filter(p_val_adj < 0.05) %>%
        pull(gene) %>%
        unique()
      
      if(length(exhaustion_t) > 0) {
        signature_definitions$T_cell_Exhaustion <- exhaustion_t
      }
    }
  }
  
  # Tissue Damage/Stress Signature
  damage_candidates <- c("HAVCR1", "KIM1", "VCAM1", "ICAM1", "LCN2", "NGAL",
                         "CXCL14", "SPP1", "VEGFA", "MT1X", "MT2A",
                         "FOS", "JUN", "EGR1", "ATF3", "DDIT3")
  
  damage_available <- intersect(damage_candidates,
                                ici_specific_cosmx %>% 
                                  filter(in_CosMx == TRUE) %>% 
                                  pull(gene))
  
  if(length(damage_available) >= 3) {
    signature_definitions$Tissue_Damage <- damage_available
  }
  
  # ========================================
  # Strategy 3: Top ICI-specific (General)
  # ========================================
  
  cat("\n--- Strategy 3: General ICI Signature ---\n")
  
  if(exists("ici_specific_cosmx")) {
    ici_cosmx_genes <- ici_specific_cosmx %>%
      filter(in_CosMx == TRUE) %>%
      arrange(desc(mean_log2FC))
    
    # Top 15 for robust signature
    signature_definitions$ICI_General_Top15 <- head(ici_cosmx_genes, 15)$gene
    
    # Known markers (if available)
    known_markers <- c("SPP1", "SOD2", "SERPINA1", "GPX3", "VEGFA", 
                       "CXCL14", "AZGP1", "G0S2", "MT1X", "AQP3")
    known_available <- intersect(known_markers, ici_cosmx_genes$gene)
    
    if(length(known_available) >= 5) {
      signature_definitions$Known_ICI_Markers <- known_available
    }
  }
  
  # ========================================
  # Print and Save Results
  # ========================================
  
  cat("\n=== Signature Summary ===\n")
  for(sig_name in names(signature_definitions)) {
    genes <- signature_definitions[[sig_name]]
    if(length(genes) > 0) {
      cat("\n", sig_name, "(", length(genes), "genes):\n")
      cat("  ", paste(genes, collapse = ", "), "\n")
    }
  }
  
  # Save signatures
  saveRDS(signature_definitions, 
          file.path(output_dir, "ICI_signature_candidates_for_CosMx.rds"))
  
  # Save as CSV with metadata
  sig_df_list <- list()
  for(sig_name in names(signature_definitions)) {
    genes <- signature_definitions[[sig_name]]
    if(length(genes) > 0) {
      # Get fold changes if available
      gene_info <- ici_specific_cosmx %>%
        filter(gene %in% genes) %>%
        select(gene, mean_log2FC, log2FC_vs_Rej, log2FC_vs_Drug)
      
      if(nrow(gene_info) > 0) {
        gene_info$Signature <- sig_name
        sig_df_list[[sig_name]] <- gene_info
      } else {
        # Genes not in ICI-specific list
        sig_df_list[[sig_name]] <- data.frame(
          gene = genes,
          mean_log2FC = NA,
          log2FC_vs_Rej = NA,
          log2FC_vs_Drug = NA,
          Signature = sig_name
        )
      }
    }
  }
  
  sig_df_complete <- bind_rows(sig_df_list)
  
  write.csv(sig_df_complete, 
            file.path(output_dir, "ICI_signature_candidates_for_CosMx_detailed.csv"),
            row.names = FALSE)
  
  # Simple version (original format)
  sig_df_simple <- data.frame(
    Signature = rep(names(signature_definitions), 
                    sapply(signature_definitions, length)),
    Gene = unlist(signature_definitions)
  )
  
  write.csv(sig_df_simple, 
            file.path(output_dir, "ICI_signature_candidates_for_CosMx.csv"),
            row.names = FALSE)
  
  cat("\nâœ… Signature candidates created and saved\n")
  cat("Total signatures:", length(signature_definitions), "\n")
  cat("Total unique genes:", length(unique(unlist(signature_definitions))), "\n\n")
}
```

---

# 5. çµæžœã‚µãƒžãƒªãƒ¼ã¨ãƒ¬ãƒãƒ¼ãƒˆ {.tabset .tabset-fade}

## ç·åˆã‚µãƒžãƒªãƒ¼

```{r comprehensive_summary}
cat("=== Comprehensive Analysis Summary ===\n")
cat(strrep("=", 70), "\n\n")

# Sample information
cat("1. SAMPLE INFORMATION\n")
cat("   Total cells analyzed:", ncol(scffpe), "\n")
cat("   Sample groups:\n")
group_summary <- table(scffpe$group)
for(g in names(group_summary)) {
  cat("     -", g, ":", group_summary[g], "cells\n")
}

# All cells DEG
cat("\n2. ALL CELLS DEG ANALYSIS\n")
if(exists("deg_ici_vs_rej_sig")) {
  cat("   ICI_AKI vs Rejection:\n")
  cat("     - Total significant genes:", nrow(deg_ici_vs_rej_sig), "\n")
  cat("     - ICI_high:", sum(deg_ici_vs_rej_sig$direction == "ICI_high"), "\n")
  cat("     - Rejection_high:", sum(deg_ici_vs_rej_sig$direction == "Rejection_high"), "\n")
}
if(exists("deg_ici_vs_drug_sig")) {
  cat("   ICI_AKI vs Drug-AKI:\n")
  cat("     - Total significant genes:", nrow(deg_ici_vs_drug_sig), "\n")
  cat("     - ICI_high:", sum(deg_ici_vs_drug_sig$direction == "ICI_high"), "\n")
  cat("     - Drug_high:", sum(deg_ici_vs_drug_sig$direction == "Drug_high"), "\n")
}
if(exists("ici_specific")) {
  cat("   ICI-specific genes (both comparisons):", nrow(ici_specific), "\n")
}

# Cell type-specific DEG
cat("\n3. CELL TYPE-SPECIFIC DEG ANALYSIS\n")
if(length(celltype_deg_results) > 0) {
  cat("   Analyzed cell types:", length(celltype_deg_results), "\n")
  cat("   Total DEG combinations:", nrow(all_celltype_deg), "\n")
}

# CosMx overlap
cat("\n4. COSMX PANEL OVERLAP\n")
if(!is.null(cosmx_genes)) {
  if(exists("ici_specific_cosmx")) {
    cat("   ICI-specific genes in CosMx:", sum(ici_specific_cosmx$in_CosMx), "/", 
        nrow(ici_specific_cosmx), 
        sprintf(" (%.1f%%)\n", sum(ici_specific_cosmx$in_CosMx)/nrow(ici_specific_cosmx)*100))
  }
  if(exists("signature_definitions")) {
    cat("   Signature candidates created:", length(signature_definitions), "\n")
    for(sig in names(signature_definitions)) {
      cat("     -", sig, ":", length(signature_definitions[[sig]]), "genes\n")
    }
  }
}

cat("\n5. OUTPUT FILES\n")
output_files <- list.files(output_dir, full.names = FALSE)
cat("   Total files created:", length(output_files), "\n")
cat("   Location:", output_dir, "\n")

cat("\n", strrep("=", 70), "\n")
cat("âœ… Analysis Complete!\n")
cat(strrep("=", 70), "\n\n")
```

## æ¬¡ã®ã‚¹ãƒ†ãƒƒãƒ—æŽ¨å¥¨

```{r next_steps}
cat("=== Recommended Next Steps ===\n\n")

cat("ðŸ“Š COMPLETED IN THIS ANALYSIS:\n")
cat("  âœ… Comprehensive DEG analysis (all cells + cell type-specific)\n")
cat("  âœ… ICI-_KI specific gene identification\n")
cat("  âœ… CosMx 1K panel overlap confirmation\n")
cat("  âœ… Signature candidates for spatial validation\n\n")

cat("ðŸŽ¯ NEXT ANALYSIS PHASE:\n")
cat("  1. CosMx Signature Projection\n")
cat("     - Load CosMx annotated data (20251117_CosMx_comprehensive_annotation.rds)\n")
cat("     - Project ICI-signatures using AddModuleScore\n")
cat("     - Visualize spatial distribution\n\n")

cat("  2. Spatial Validation\n")
cat("     - Density comparison (ICI vs non-ICI)\n")
cat("     - Cell-cell interaction analysis\n")
cat("     - Hotspot identification\n\n")

cat("  3. Pathway Enrichment\n")
cat("     - GO/KEGG analysis on DEG lists\n")
cat("     - Functional annotation\n")
cat("     - Network analysis\n\n")

cat("ðŸ“ KEY OUTPUT FILES TO USE:\n")
cat("  - ICI_specific_genes_CosMx_overlap.csv\n")
cat("  - ICI_signature_candidates_for_CosMx.rds\n")
cat("  - celltype_specific_DEG_results.rds\n")
cat("  - celltype_DEG_CosMx_overlap.rds\n\n")

cat("ðŸ’¡ TIPS FOR SPATIAL PROJECTION:\n")
cat("  - Use 5-10 genes per signature for robust module scores\n")
cat("  - Validate with known markers (e.g., SPP1)\n")
cat("  - Compare signature scores across conditions\n")
cat("  - Focus on cell types with good marker coverage\n\n")
```

## ã‚»ãƒƒã‚·ãƒ§ãƒ³æƒ…å ±

```{r session_info}
cat("=== Session Information ===\n")
sessionInfo()
```

---

# ã¾ã¨ã‚

## ä¸»è¦ãªæˆæžœ

1. **åŒ…æ‹¬çš„DEGè§£æžã®å®Œäº†**
   - å…¨ç´°èƒžãƒ¬ãƒ™ãƒ«ã§ã®3ç¾¤æ¯”è¼ƒ
   - ç´°èƒžã‚¿ã‚¤ãƒ—åˆ¥ã®è©³ç´°è§£æž
   - ICI-AKIç‰¹ç•°çš„éºä¼å­ã®åŒå®š

2. **CosMxçµ±åˆæº–å‚™ã®å®Œäº†**
   - 1Kãƒ‘ãƒãƒ«ã¨ã®é‡è¤‡ç¢ºèª
   - Signatureå€™è£œã®ä½œæˆ
   - ç©ºé–“æŠ•å½±ç”¨ãƒ‡ãƒ¼ã‚¿ã®æ•´å‚™

3. **æ¬¡ãƒ•ã‚§ãƒ¼ã‚ºã¸ã®æº–å‚™**
   - æ˜Žç¢ºãªå‡ºåŠ›ãƒ•ã‚¡ã‚¤ãƒ«
   - å†ç¾å¯èƒ½ãªè§£æžãƒ•ãƒ­ãƒ¼
   - è©³ç´°ãªãƒ‰ã‚­ãƒ¥ãƒ¡ãƒ³ãƒ†ãƒ¼ã‚·ãƒ§ãƒ³

## æŠ€è¡“çš„ç‰¹å¾´

- **é ‘å¥ãªçµ±è¨ˆæ‰‹æ³•**: Wilcoxonæ¤œå®š + BHè£œæ­£
- **ç”Ÿç‰©å­¦çš„æ„ç¾©é‡è¦–**: Effect size (log2FC) + çµ±è¨ˆçš„æœ‰æ„æ€§
- **å®Ÿç”¨çš„ãªå‡ºåŠ›**: CosMxæŠ•å½±å¯èƒ½ãªãƒ•ã‚©ãƒ¼ãƒžãƒƒãƒˆ

## å­¦è¡“çš„ä¾¡å€¤

- ICI-AKIç‰¹ç•°çš„ãªåˆ†å­ã‚·ã‚°ãƒãƒãƒ£ã®ç™ºè¦‹
- ç´°èƒžã‚¿ã‚¤ãƒ—åˆ¥ã®ç—…æ…‹ãƒ¡ã‚«ãƒ‹ã‚ºãƒ è§£æ˜Ž
- ç©ºé–“ãƒˆãƒ©ãƒ³ã‚¹ã‚¯ãƒªãƒ—ãƒˆãƒŸã‚¯ã‚¹ã¨ã®çµ±åˆåŸºç›¤

---

**Analysis completed**: `r Sys.time()`

**Output directory**: `r output_dir`
