---
title: "CosMx Comprehensive Cell Type Annotation from Scratch"
author: "Kidney ICI-AKI Analysis Project"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
    code_folding: show
    theme: cosmo
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 14,
  fig.height = 10,
  dpi = 300
)
```

# ãƒ—ãƒ­ã‚¸ã‚§ã‚¯ãƒˆæ¦‚è¦

**ç›®çš„**: CosMx 1,000éºä¼å­ãƒ‘ãƒãƒ«ã‚’æ´»ç”¨ã—ãŸè…Žè‡“çµ„ç¹”ã®åŒ…æ‹¬çš„ç´°èƒžã‚¿ã‚¤ãƒ—ã‚¢ãƒŽãƒ†ãƒ¼ã‚·ãƒ§ãƒ³

**å…¥åŠ›ãƒ‡ãƒ¼ã‚¿**: 
- File: `20251006 cosmx_sctransform_harmony_merged.rds`
- Cells: 319,141ç´°èƒž
- Samples: S0 (Rejection), S1/S3 (ICI-AKI), S2 (Drug-AKI)
- Normalization: SCTransform + Harmony batch correction

**ã‚¢ãƒ—ãƒ­ãƒ¼ãƒ**: 
1. CosMx Universal Cell Characterization Panel (1,000éºä¼å­)ã®å®Œå…¨æ´»ç”¨
2. Module score-basedéšŽå±¤çš„åˆ†é¡ž
3. æ—¢çŸ¥ãƒžãƒ¼ã‚«ãƒ¼éºä¼å­ã«ã‚ˆã‚‹æ¤œè¨¼
4. ç©ºé–“æƒ…å ±ã‚’ç”¨ã„ãŸå¦¥å½“æ€§ç¢ºèª

**æœŸå¾…ã•ã‚Œã‚‹æˆæžœ**:
- Major cell types: 5-8ã‚¿ã‚¤ãƒ—
- Detailed cell types: 20-30ã‚¿ã‚¤ãƒ—
- ä¿¡é ¼æ€§ã®é«˜ã„ã‚¢ãƒŽãƒ†ãƒ¼ã‚·ãƒ§ãƒ³ï¼ˆç©ºé–“æ¤œè¨¼æ¸ˆã¿ï¼‰

---

# 1. ç’°å¢ƒæº–å‚™ {.tabset .tabset-fade}

## ãƒ©ã‚¤ãƒ–ãƒ©ãƒªèª­ã¿è¾¼ã¿

```{r load_libraries}
cat("=== Loading Required Libraries ===\n")

suppressPackageStartupMessages({
  library(Seurat)
  library(SeuratObject)
  library(dplyr)
  library(tidyr)
  library(ggplot2)
  library(patchwork)
  library(viridis)
  library(RColorBrewer)
  library(pheatmap)
  library(readxl)
})

cat("âœ… All libraries loaded\n")
cat("Seurat version:", as.character(packageVersion("Seurat")), "\n")
cat("R version:", R.version.string, "\n\n")
```

## ãƒ‡ã‚£ãƒ¬ã‚¯ãƒˆãƒªè¨­å®š

```{r setup_directories}
cat("=== Setting up Directories ===\n")

# Project root
project_root <- "/Users/HiroyukiArai/Documents/2025å¹´ç•™å­¦/ç ”ç©¶é–¢é€£/ç ”ç©¶çµæžœ/202509 cosmx-kidney-project"

# Data directory
data_dir <- file.path(project_root, "data/processed")

# Output directory
output_dir <- file.path(project_root, "results/figures/20251117_Comprehensive_Annotation")
dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)

# Gene panel file
panel_file <- "/Users/HiroyukiArai/Documents/å¤§å­¦é™¢/ç ”ç©¶çµæžœ/2024æ‘ä¸Šå…ˆç”ŸCosMX/CosMxâ„¢ Human Universal Cell Characterization Panel Gene Target List.xlsx"

cat("Data directory:", data_dir, "\n")
cat("Output directory:", output_dir, "\n")
cat("Panel file:", panel_file, "\n\n")
```

## ãƒ‡ãƒ¼ã‚¿èª­ã¿è¾¼ã¿

```{r load_data}
cat("=== Loading CosMx Data ===\n")

# Load SCTransform + Harmony integrated data
cosmx_file <- file.path(data_dir, "20251006 cosmx_sctransform_harmony_merged.rds")

if(!file.exists(cosmx_file)) {
  stop("âŒ Data file not found: ", cosmx_file)
}

cat("Loading data (this may take a few minutes)...\n")
cosmx <- readRDS(cosmx_file)

cat("âœ… Data loaded successfully\n")
cat("Total cells:", ncol(cosmx), "\n")
cat("Total genes:", nrow(cosmx), "\n")
cat("Samples:", paste(unique(cosmx$sample), collapse = ", "), "\n")
cat("Assays:", names(cosmx@assays), "\n")
cat("Reductions:", names(cosmx@reductions), "\n\n")

# Set default assay
DefaultAssay(cosmx) <- "SCT"
cat("âœ… Default assay set to SCT\n\n")
```

## CosMxãƒ‘ãƒãƒ«éºä¼å­ãƒªã‚¹ãƒˆèª­ã¿è¾¼ã¿

```{r load_cosmx_panel}
cat("=== Loading CosMx Gene Panel ===\n")

# Read gene list from Annotations sheet
cosmx_panel_df <- read_excel(panel_file, sheet = "Annotations", skip = 1) 

# Extract gene names (first column after header)
cosmx_genes_panel <- cosmx_panel_df[[1]]  # Header already skipped with skip=1
cosmx_genes_panel <- cosmx_genes_panel[!is.na(cosmx_genes_panel)]
cosmx_genes_panel <- cosmx_genes_panel[cosmx_genes_panel != "Gene"]

cat("âœ… CosMx panel genes loaded\n")
cat("Total panel genes:", length(cosmx_genes_panel), "\n")

# Check overlap with our data
genes_in_data <- intersect(cosmx_genes_panel, rownames(cosmx))
cat("Genes detected in data:", length(genes_in_data), 
    sprintf("(%.1f%%)\n", length(genes_in_data)/length(cosmx_genes_panel)*100))

# Load functional annotations
cosmx_annotations <- read_excel(panel_file, sheet = "Annotations", skip = 1)
colnames(cosmx_annotations)[1] <- "Gene"


# Extract key functional categories
functional_categories <- colnames(cosmx_annotations)[4:ncol(cosmx_annotations)]
cat("Functional categories:", length(functional_categories), "\n")
cat("Examples:", paste(head(functional_categories, 10), collapse = ", "), "...\n\n")
```

---

# 2. CosMxãƒ‘ãƒãƒ«è…Žè‡“ç‰¹ç•°çš„ãƒžãƒ¼ã‚«ãƒ¼å®šç¾© {.tabset .tabset-fade}

## ä¸Šçš®ç´°èƒžãƒžãƒ¼ã‚«ãƒ¼ - åˆæœŸå®šç¾©ã¨åˆ¶ç´„ã®ç¢ºèª
```{r epithelial_markers_initial_check, fig.width=12, fig.height=6}
cat("=== Defining Epithelial Cell Markers - Initial Attempt ===\n")
cat("Testing classical nephron segment markers...\n\n")

# å¤å…¸çš„ãªè…Žè‡“ä¸Šçš®ãƒžãƒ¼ã‚«ãƒ¼ï¼ˆç†æƒ³çš„ãªãƒªã‚¹ãƒˆï¼‰
epithelial_markers_classical <- list(
  # General epithelial
  Epithelial_General = c("EPCAM", "KRT18", "KRT8", "KRT19", "CDH1"),
  
  # Proximal tubule
  Proximal_Tubule = c("LRP2", "SLC34A1", "SLC5A2", "CUBN", "SLC22A6", 
                      "SLC22A8", "HAVCR1"),
  
  # Distal tubule
  Distal_Tubule = c("SLC12A3", "CALB1", "TRPM6", "PVALB"),
  
  # Thick ascending limb
  Thick_Ascending_Limb = c("SLC12A1", "UMOD", "CLDN10", "CLDN16"),
  
  # Collecting duct
  Collecting_Duct = c("AQP2", "AQP3", "AQP4", "ATP6V1B1", "ATP6V0D2", 
                      "SLC26A4", "SCNN1A", "SCNN1B", "SCNN1G"),
  
  # Podocyte (Glomerulus)
  Podocyte = c("NPHS1", "NPHS2", "PODXL", "WT1", "SYNPO", "MAGI2"),
  
  # Parietal epithelial cell
  Parietal_Epithelial = c("PAX8", "CDH1", "CLDN1")
)

# Check availability in data
epithelial_availability <- data.frame()
epithelial_available <- list()

for(cell_type in names(epithelial_markers_classical)) {
  genes <- epithelial_markers_classical[[cell_type]]
  available <- intersect(genes, rownames(cosmx))
  epithelial_available[[cell_type]] <- available
  
  epithelial_availability <- rbind(epithelial_availability, data.frame(
    Cell_Type = cell_type,
    Total_Markers = length(genes),
    Available_Markers = length(available),
    Availability_Pct = round(length(available) / length(genes) * 100, 1),
    Available_Genes = paste(available, collapse = ", ")
  ))
  
  cat(sprintf("%-25s: %2d/%2d genes (%.1f%%) - %s\n", 
              cell_type, 
              length(available), 
              length(genes),
              length(available) / length(genes) * 100,
              ifelse(length(available) > 0, 
                     paste(available, collapse = ", "), 
                     "None")))
}

cat("\n")

# Save availability table
write.csv(epithelial_availability, 
          file.path(output_dir, "epithelial_marker_availability.csv"),
          row.names = FALSE)

# Visualization of marker availability
library(ggplot2)

# Create summary for plotting
epi_summary <- epithelial_availability %>%
  select(Cell_Type, Available_Markers, Total_Markers) %>%
  mutate(
    Missing_Markers = Total_Markers - Available_Markers,
    Availability_Status = case_when(
      Available_Markers == 0 ~ "No markers (0%)",
      Available_Markers == 1 ~ "Single marker only",
      Available_Markers >= 2 & Available_Markers < Total_Markers ~ "Partial",
      Available_Markers == Total_Markers ~ "Complete"
    )
  )

# Stacked bar plot
p_availability <- ggplot(epi_summary, 
                         aes(x = reorder(Cell_Type, Available_Markers))) +
  geom_bar(aes(y = Available_Markers, fill = "Available"), 
           stat = "identity", alpha = 0.7) +
  geom_bar(aes(y = Missing_Markers), 
           stat = "identity", fill = "grey70", alpha = 0.5,
           position = position_stack(vjust = 1, reverse = FALSE)) +
  geom_text(aes(y = Total_Markers, 
                label = sprintf("%d/%d", Available_Markers, Total_Markers)),
            hjust = -0.2, size = 3.5, fontface = "bold") +
  coord_flip() +
  scale_fill_manual(values = c("Available" = "#2E7D32")) +
  labs(
    title = "Epithelial Marker Availability in CosMx Panel",
    subtitle = "Classical nephron segment markers are severely limited",
    x = "Cell Type",
    y = "Number of Markers",
    fill = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0),
    plot.subtitle = element_text(color = "red", hjust = 0),
    legend.position = "bottom",
    panel.grid.major.y = element_blank()
  ) +
  ylim(0, max(epi_summary$Total_Markers) * 1.15)

print(p_availability)

ggsave(
  file.path(output_dir, "epithelial_marker_availability.png"),
  p_availability, width = 10, height = 6, dpi = 300
)

cat("âœ… Marker availability plot saved\n\n")
```

```{r epithelial_markers_limitation_summary}
cat("=== CRITICAL FINDING: Insufficient Nephron Segment Markers ===\n\n")

# Identify problematic subtypes
insufficient_subtypes <- epithelial_availability %>%
  filter(Available_Markers < 2) %>%
  arrange(Available_Markers)

cat("âš ï¸  Cell types with insufficient markers (< 2 genes):\n")
for(i in 1:nrow(insufficient_subtypes)) {
  cat(sprintf("   %d. %-25s: %d/%d markers (%.0f%%)\n",
              i,
              insufficient_subtypes$Cell_Type[i],
              insufficient_subtypes$Available_Markers[i],
              insufficient_subtypes$Total_Markers[i],
              insufficient_subtypes$Availability_Pct[i]))
}

cat("\nðŸ“Š Summary statistics:\n")
cat(sprintf("   - Total epithelial subtypes: %d\n", nrow(epithelial_availability)))
cat(sprintf("   - Subtypes with 0 markers: %d\n", 
            sum(epithelial_availability$Available_Markers == 0)))
cat(sprintf("   - Subtypes with 1 marker: %d\n", 
            sum(epithelial_availability$Available_Markers == 1)))
cat(sprintf("   - Subtypes with â‰¥2 markers: %d\n", 
            sum(epithelial_availability$Available_Markers >= 2)))

cat("\nâš ï¸  CONCLUSION:\n")
cat("   Classical nephron segment annotation (PT/TAL/DCT/CD/Podocyte) is\n")
cat("   NOT FEASIBLE with the current CosMx 1K panel.\n\n")

cat("âž¡ï¸  STRATEGY CHANGE:\n")
cat("   1. Use broad 'Epithelial' category based on general markers\n")
cat("   2. Reference markers (CALB1, AQP3) for visualization only\n")
cat("   3. Future: Add cell STATE signatures from scFFPE analysis\n\n")
```

```{r epithelial_markers_revised_definition}
cat("=== Revised Epithelial Marker Definition (Realistic Approach) ===\n\n")

# Only use markers with sufficient coverage
epithelial_markers_revised <- list(
  # This is reliable - 5/5 markers available
  Epithelial_General = epithelial_available$Epithelial_General
)

# Add reference markers (for visualization, not classification)
reference_markers <- list()
if(length(epithelial_available$Distal_Tubule) > 0) {
  reference_markers$Reference_Distal <- epithelial_available$Distal_Tubule
}
if(length(epithelial_available$Collecting_Duct) > 0) {
  reference_markers$Reference_Collecting <- epithelial_available$Collecting_Duct
}

cat("âœ… MARKERS FOR CLASSIFICATION:\n")
for(marker_type in names(epithelial_markers_revised)) {
  genes <- epithelial_markers_revised[[marker_type]]
  cat(sprintf("   %-25s: %d genes (%s)\n",
              marker_type,
              length(genes),
              paste(genes, collapse = ", ")))
}

cat("\nðŸ“‹ MARKERS FOR REFERENCE/VISUALIZATION ONLY:\n")
if(length(reference_markers) > 0) {
  for(marker_type in names(reference_markers)) {
    genes <- reference_markers[[marker_type]]
    cat(sprintf("   %-25s: %d genes (%s)\n",
                marker_type,
                length(genes),
                paste(genes, collapse = ", ")))
  }
} else {
  cat("   (None available)\n")
}

cat("\nðŸ’¡ RATIONALE:\n")
cat("   Module scores require â‰¥2 genes for robustness.\n")
cat("   Single-gene markers (CALB1, AQP3) are shown for reference\n")
cat("   but NOT used for formal cell type classification.\n\n")

cat("ðŸ“ FOR METHODS SECTION:\n")
cat('   "Because the CosMx 1K panel contains only limited nephron\n')
cat('    segment markers (0/7 PT, 0/4 TAL, 1/9 CD, 0/6 Podocyte),\n')
cat('    we did not attempt precise nephron segment annotation and\n')
cat('    instead treated tubular epithelial cells as a broad category."\n\n')
```

## å…ç–«ç´°èƒžãƒžãƒ¼ã‚«ãƒ¼

```{r immune_markers}
cat("=== Defining Immune Cell Markers ===\n")

# Immune cell markers based on CosMx panel
immune_markers <- list(
  # Pan-immune
  Immune_General = c("PTPRC"),  # CD45
  
  # T cells
  T_cell_General = c("CD3D", "CD3E", "CD3G", "CD2", "CD7", "CD5"),
  CD4_T_cell = c("CD4", "IL7R", "LDHB"),
  CD8_T_cell = c("CD8A", "CD8B", "GZMA", "GZMB", "PRF1", "NKG7"),
  Treg = c("FOXP3", "IL2RA", "CTLA4", "IKZF2", "TIGIT"),
  
  # B cells
  B_cell = c("CD19", "MS4A1", "CD79A", "CD79B", "CD22", "CD24"),
  Plasma_cell = c("SDC1", "MZB1", "JCHAIN", "IGHG1", "IGKC", "IGHM", "XBP1"),
  
  # Myeloid cells
  Macrophage = c("CD68", "CD163", "MSR1", "MRC1", "CD14", "FCGR1A", 
                 "CSF1R", "ITGAM", "CD86", "CD80"),
  Monocyte = c("CD14", "FCGR3A", "S100A8", "S100A9", "S100A12", "LYZ"),
  Dendritic_cell = c("ITGAX", "CD1C", "FCER1A", "CLEC9A", "CLEC10A", 
                     "LAMP3", "CD1A", "CD1E"),
  
  # NK cells
  NK_cell = c("NCAM1", "NKG7", "GNLY", "PRF1", "GZMA", "GZMB", 
              "KLRB1", "KLRD1", "KLRC1", "KLRC2", "FCGR3A"),
  
  # NK T cells
  NK_T_cell = c("KLRB1", "CD3E", "CD3D", "NKG7")
)

# Check availability in data
immune_available <- list()
for(cell_type in names(immune_markers)) {
  genes <- immune_markers[[cell_type]]
  available <- intersect(genes, rownames(cosmx))
  immune_available[[cell_type]] <- available
  
  cat(sprintf("%-25s: %2d/%2d genes (%s)\n", 
              cell_type, 
              length(available), 
              length(genes),
              ifelse(length(available) > 0, 
                     paste(head(available, 3), collapse=", "),
                     "None")))
}

cat("\n")
```

## é–“è³ªãƒ»è¡€ç®¡ç³»ãƒžãƒ¼ã‚«ãƒ¼

```{r stromal_vascular_markers}
cat("=== Defining Stromal & Vascular Markers ===\n")

# Stromal and vascular markers
stromal_vascular_markers <- list(
  # Endothelial cells
  Endothelial = c("PECAM1", "VWF", "CDH5", "FLT1", "KDR", "TEK", 
                  "CLDN5", "CD34", "ENG"),
  
  # Fibroblasts
  Fibroblast = c("COL1A1", "COL1A2", "COL3A1", "DCN", "LUM", "PDGFRA",
                 "VIM", "FN1", "ACTA2"),
  
  # Myofibroblasts
  Myofibroblast = c("ACTA2", "TAGLN", "MYL9", "CNN1", "PDGFRB"),
  
  # Mesangial cells
  Mesangial = c("PDGFRB", "ACTA2", "REN", "DES", "ITGA8"),
  
  # Pericytes
  Pericyte = c("PDGFRB", "RGS5", "NOTCH3", "CSPG4", "MCAM"),
  
  # Smooth muscle cells
  Smooth_Muscle = c("ACTA2", "MYH11", "TAGLN", "CNN1", "MYLK")
)

# Check availability in data
stromal_available <- list()
for(cell_type in names(stromal_vascular_markers)) {
  genes <- stromal_vascular_markers[[cell_type]]
  available <- intersect(genes, rownames(cosmx))
  stromal_available[[cell_type]] <- available
  
  cat(sprintf("%-25s: %2d/%2d genes (%s)\n", 
              cell_type, 
              length(available), 
              length(genes),
              paste(available, collapse=", ")))
}

cat("\n")
```

---

# 3. Module Scoreè¨ˆç®— {.tabset .tabset-fade}

## ä¸Šçš®ç´°èƒžã‚¹ã‚³ã‚¢

```{r epithelial_module_scores_realistic}
cat("=== Calculating Epithelial Cell Module Scores (Realistic) ===\n")

# Epithelial_General ã®ã¿è¨ˆç®—ï¼ˆã“ã‚Œã¯ä¿¡é ¼ã§ãã‚‹ï¼‰
if("Epithelial_General" %in% names(epithelial_markers_revised)) {
  genes <- epithelial_markers_revised$Epithelial_General
  
  if(length(genes) >= 2) {
    cosmx <- AddModuleScore(
      cosmx,
      features = list(genes),
      assay = "SCT",
      name = "Epithelial_General_Score",
      ctrl = 10,
      nbin = 10
    )
    
    # Rename to remove number suffix
    score_col <- "Epithelial_General_Score1"
    if(score_col %in% colnames(cosmx@meta.data)) {
      colnames(cosmx@meta.data)[colnames(cosmx@meta.data) == score_col] <- 
        "Epithelial_General_Score"
    }
    
    cat(sprintf("âœ… Epithelial_General_Score: %d genes (%s)\n", 
                length(genes),
                paste(genes, collapse = ", ")))
  } else {
    cat("âŒ Insufficient genes for Epithelial_General\n")
  }
}

cat("\nâ­ï¸  Skipping nephron segment subtypes (insufficient markers)\n")
cat("   Proximal_Tubule: 0/7 markers\n")
cat("   Thick_Ascending_Limb: 0/4 markers\n")
cat("   Distal_Tubule: 1/4 markers (insufficient)\n")
cat("   Collecting_Duct: 1/9 markers (insufficient)\n")
cat("   Podocyte: 0/6 markers\n")
cat("\nâž¡ï¸  Future: Will add cell STATE signatures from scFFPE analysis\n\n")
```

## å…ç–«ç´°èƒžã‚¹ã‚³ã‚¢

```{r immune_module_scores}
cat("=== Calculating Immune Cell Module Scores ===\n")

# Calculate module scores for immune cell types
for(cell_type in names(immune_available)) {
  genes <- immune_available[[cell_type]]
  
  if(length(genes) >= 2) {
    score_name <- paste0(cell_type, "_Score")
    
    cosmx <- AddModuleScore(
      cosmx,
      features = list(genes),
      assay = "SCT",
      name = score_name,
      ctrl = 10,
      nbin = 10
    )
    
    # Rename to remove number suffix
    score_col <- paste0(score_name, "1")
    if(score_col %in% colnames(cosmx@meta.data)) {
      colnames(cosmx@meta.data)[colnames(cosmx@meta.data) == score_col] <- score_name
    }
    
    cat(sprintf("âœ… %-25s: %d genes\n", cell_type, length(genes)))
  } else {
    cat(sprintf("â­ï¸  %-25s: Skipped (insufficient genes: %d)\n", 
                cell_type, length(genes)))
  }
}

cat("\n")
```

## é–“è³ªãƒ»è¡€ç®¡ç³»ã‚¹ã‚³ã‚¢

```{r stromal_module_scores}
cat("=== Calculating Stromal & Vascular Module Scores ===\n")

# Calculate module scores for stromal/vascular cell types
for(cell_type in names(stromal_available)) {
  genes <- stromal_available[[cell_type]]
  
  if(length(genes) >= 2) {
    score_name <- paste0(cell_type, "_Score")
    
    cosmx <- AddModuleScore(
      cosmx,
      features = list(genes),
      assay = "SCT",
      name = score_name,
      ctrl = 10,
      nbin = 10
    )
    
    # Rename to remove number suffix
    score_col <- paste0(score_name, "1")
    if(score_col %in% colnames(cosmx@meta.data)) {
      colnames(cosmx@meta.data)[colnames(cosmx@meta.data) == score_col] <- score_name
    }
    
    cat(sprintf("âœ… %-25s: %d genes\n", cell_type, length(genes)))
  } else {
    cat(sprintf("â­ï¸  %-25s: Skipped (insufficient genes: %d)\n", 
                cell_type, length(genes)))
  }
}

cat("\n")
```

## ã‚¹ã‚³ã‚¢ã‚µãƒžãƒªãƒ¼

```{r score_summary}
cat("=== Module Score Summary ===\n")

# Get all score columns
score_cols <- grep("_Score$", colnames(cosmx@meta.data), value = TRUE)

cat("Total module scores calculated:", length(score_cols), "\n\n")

# Summary statistics for each score
score_summary <- data.frame()

for(score in score_cols) {
  cell_type <- gsub("_Score$", "", score)
  values <- cosmx@meta.data[[score]]
  
  score_summary <- rbind(score_summary, data.frame(
    Cell_Type = cell_type,
    Mean = round(mean(values, na.rm = TRUE), 4),
    Median = round(median(values, na.rm = TRUE), 4),
    SD = round(sd(values, na.rm = TRUE), 4),
    Min = round(min(values, na.rm = TRUE), 4),
    Max = round(max(values, na.rm = TRUE), 4),
    Q90 = round(quantile(values, 0.9, na.rm = TRUE), 4)
  ))
}

print(score_summary)

# Save summary
write.csv(score_summary, 
          file.path(output_dir, "module_score_summary.csv"),
          row.names = FALSE)

cat("\nâœ… Module score calculation complete\n\n")
```

---

# 4. éšŽå±¤çš„ç´°èƒžã‚¿ã‚¤ãƒ—åˆ†é¡ž {.tabset .tabset-fade}

## Level 1: Major Cell Types

## Level 1: Major Cell Typesï¼ˆä¿®æ­£ç‰ˆï¼‰
```{r major_cell_classification_realistic}
cat("=== Level 1: Major Cell Type Classification (Realistic) ===\n")

# Initialize
cosmx$major_celltype <- "Unclassified"

# 1. CD45+ Immune cells (åŽ³å¯†åˆ¤å®š)
DefaultAssay(cosmx) <- "SCT"
ptprc_expr <- GetAssayData(cosmx, assay = "SCT", layer = "data")["PTPRC", ]
immune_cells <- ptprc_expr > 0

cosmx$major_celltype[immune_cells] <- "Immune"
cat(sprintf("âœ… Immune: %d cells (%.2f%%)\n", 
            sum(immune_cells), 
            sum(immune_cells)/ncol(cosmx)*100))

# 2. Endothelial cells
if("Endothelial_Score" %in% colnames(cosmx@meta.data)) {
  endo_threshold <- quantile(cosmx$Endothelial_Score, 0.95)
  endo_cells <- cosmx$Endothelial_Score > endo_threshold & 
                cosmx$major_celltype == "Unclassified"
  
  cosmx$major_celltype[endo_cells] <- "Endothelial"
  cat(sprintf("âœ… Endothelial: %d cells (%.2f%%)\n", 
              sum(endo_cells), 
              sum(endo_cells)/ncol(cosmx)*100))
}

# 3. Fibroblast/Stromal cells
if("Fibroblast_Score" %in% colnames(cosmx@meta.data)) {
  fibro_threshold <- quantile(cosmx$Fibroblast_Score, 0.90)
  fibro_cells <- cosmx$Fibroblast_Score > fibro_threshold & 
                 cosmx$major_celltype == "Unclassified"
  
  cosmx$major_celltype[fibro_cells] <- "Stromal"
  cat(sprintf("âœ… Stromal: %d cells (%.2f%%)\n", 
              sum(fibro_cells), 
              sum(fibro_cells)/ncol(cosmx)*100))
}

# 4. Epithelial cells (broad category)
if("Epithelial_General_Score" %in% colnames(cosmx@meta.data)) {
  # ã‚ˆã‚Šç·©ã„é–¾å€¤ã§ä¸Šçš®ç´°èƒžã‚’åºƒãæ‹¾ã†
  epi_threshold <- quantile(
    cosmx$Epithelial_General_Score[cosmx$major_celltype == "Unclassified"], 
    0.30  # 30%ileä»¥ä¸Šã‚’ä¸Šçš®ã¨ã™ã‚‹
  )
  epi_cells <- cosmx$Epithelial_General_Score > epi_threshold & 
               cosmx$major_celltype == "Unclassified"
  
  cosmx$major_celltype[epi_cells] <- "Epithelial"
  cat(sprintf("âœ… Epithelial (broad): %d cells (%.2f%%)\n", 
              sum(epi_cells), 
              sum(epi_cells)/ncol(cosmx)*100))
}

# 5. Remaining as "Other"
other_cells <- cosmx$major_celltype == "Unclassified"
cosmx$major_celltype[other_cells] <- "Other"
cat(sprintf("âœ… Other: %d cells (%.2f%%)\n", 
            sum(other_cells), 
            sum(other_cells)/ncol(cosmx)*100))

# Summary
cat("\n=== Major Cell Type Distribution ===\n")
major_summary <- table(cosmx$major_celltype)
print(major_summary)
print(round(prop.table(major_summary) * 100, 2))

cat("\nðŸ“ Note: 'Epithelial' is a BROAD category\n")
cat("   Nephron segment annotation is NOT attempted due to marker limitations\n\n")
```

## Level 2: Immune Subtypes

```{r immune_subtype_classification}
cat("=== Level 2: Immune Cell Subtype Classification ===\n")

# Extract immune cells
immune_cells_idx <- cosmx$major_celltype == "Immune"
cosmx$immune_subtype <- NA

if(sum(immune_cells_idx) > 0) {
  
  # Get all immune score columns
  immune_score_cols <- grep("^(T_cell|B_cell|Plasma|Macrophage|Monocyte|Dendritic|NK)", 
                             colnames(cosmx@meta.data), value = TRUE)
  immune_score_cols <- grep("_Score$", immune_score_cols, value = TRUE)
  
  if(length(immune_score_cols) > 0) {
    # For each immune cell, find the highest scoring cell type
    immune_scores <- cosmx@meta.data[immune_cells_idx, immune_score_cols, drop = FALSE]
    
    # Find cell type with maximum score for each cell
    max_scores <- apply(immune_scores, 1, function(x) {
      if(all(is.na(x))) return(NA)
      names(x)[which.max(x)]
    })
    
    # Clean up names (remove _Score suffix)
    max_scores <- gsub("_Score$", "", max_scores)
    
    # Assign subtypes
    cosmx$immune_subtype[immune_cells_idx] <- max_scores
    
    # Summary
    cat("âœ… Immune subtype classification complete\n")
    immune_subtype_table <- table(cosmx$immune_subtype, useNA = "ifany")
    print(immune_subtype_table)
    
  } else {
    cat("âš ï¸ No immune score columns found\n")
  }
} else {
  cat("âš ï¸ No immune cells detected\n")
}

cat("\n")
```

## Level 2: Epithelial Subtypes (Reference markers only)
```{r epithelial_reference_markers}
cat("=== Level 2: Epithelial - Reference Markers Only ===\n")

# Extract epithelial cells
epi_cells_idx <- cosmx$major_celltype == "Epithelial"
cosmx$epithelial_subtype <- NA  # Initialize as NA (no formal subtyping)

if(sum(epi_cells_idx) > 0) {
  
  cat(sprintf("Total epithelial cells: %d (%.2f%%)\n\n", 
              sum(epi_cells_idx),
              sum(epi_cells_idx) / ncol(cosmx) * 100))
  
  # Add reference marker expression (for visualization only)
  DefaultAssay(cosmx) <- "SCT"
  
  # CALB1 expression (Distal tubule reference)
  if("CALB1" %in% rownames(cosmx)) {
    calb1_expr <- GetAssayData(cosmx, assay = "SCT", layer = "data")["CALB1", ]
    cosmx$CALB1_expression <- calb1_expr
    
    calb1_pos_epi <- sum(calb1_expr[epi_cells_idx] > 0)
    cat(sprintf("ðŸ“‹ CALB1+ epithelial cells: %d (%.1f%% of epithelial)\n", 
                calb1_pos_epi, 
                calb1_pos_epi / sum(epi_cells_idx) * 100))
    cat("   â†’ Reference for potential distal tubule cells\n\n")
  }
  
  # AQP3 expression (Collecting duct reference)
  if("AQP3" %in% rownames(cosmx)) {
    aqp3_expr <- GetAssayData(cosmx, assay = "SCT", layer = "data")["AQP3", ]
    cosmx$AQP3_expression <- aqp3_expr
    
    aqp3_pos_epi <- sum(aqp3_expr[epi_cells_idx] > 0)
    cat(sprintf("ðŸ“‹ AQP3+ epithelial cells: %d (%.1f%% of epithelial)\n", 
                aqp3_pos_epi, 
                aqp3_pos_epi / sum(epi_cells_idx) * 100))
    cat("   â†’ Reference for potential collecting duct cells\n\n")
  }
  
  cat("âš ï¸  IMPORTANT NOTES:\n")
  cat("   - These are REFERENCE markers for visualization only\n")
  cat("   - NOT used for formal cell type annotation\n")
  cat("   - Single markers are insufficient for robust classification\n")
  cat("   - All epithelial cells remain in broad 'Epithelial' category\n\n")
  
  cat("ðŸ“ FOR METHODS:\n")
  cat('   "Reference markers (CALB1, AQP3) were visualized for\n')
  cat('    exploratory purposes but not used for formal annotation\n')
  cat('    due to insufficient marker gene coverage."\n\n')
  
} else {
  cat("âš ï¸ No epithelial cells detected\n")
}
```

## Unified Detailed Cell Type (Realistic)
```{r unified_detailed_celltype_realistic}
cat("=== Creating Unified Detailed Cell Type Column ===\n")

# Start with major cell types
cosmx$detailed_celltype <- cosmx$major_celltype

# Fill in immune subtypes (ã“ã‚Œã¯OK - ãƒžãƒ¼ã‚«ãƒ¼å……å®Ÿ)
immune_idx <- !is.na(cosmx$immune_subtype)
if(sum(immune_idx) > 0) {
  cosmx$detailed_celltype[immune_idx] <- cosmx$immune_subtype[immune_idx]
  cat(sprintf("âœ… Immune subtypes assigned: %d cells\n", sum(immune_idx)))
}

# Epithelial stays as "Epithelial" (NO subtyping)
epi_idx <- cosmx$major_celltype == "Epithelial"
if(sum(epi_idx) > 0) {
  cat(sprintf("ðŸ“‹ Epithelial cells kept as broad category: %d cells\n", sum(epi_idx)))
}

# Summary
cat("\n=== Final Cell Type Distribution ===\n")
detailed_summary <- table(cosmx$detailed_celltype)
print(sort(detailed_summary, decreasing = TRUE))

cat("\nðŸ“ CLASSIFICATION STRATEGY:\n")
cat("   âœ… Immune cells: Detailed subtypes (9-11 types)\n")
cat("      - Markers well-represented in CosMx panel\n")
cat("   ðŸ“‹ Epithelial cells: Broad category only\n")
cat("      - Classical nephron markers insufficient (0-1 genes per subtype)\n")
cat("   âœ… Stromal/Endothelial: Well-classified\n")
cat("      - Markers well-represented in CosMx panel\n\n")

# Save summary
write.csv(
  data.frame(
    Cell_Type = names(detailed_summary),
    Count = as.numeric(detailed_summary),
    Percentage = round(as.numeric(detailed_summary) / ncol(cosmx) * 100, 2)
  ),
  file.path(output_dir, "realistic_celltype_distribution.csv"),
  row.names = FALSE
)

cat("âœ… Cell type annotation complete (realistic approach)\n")
cat("Total unique cell types:", length(unique(cosmx$detailed_celltype)), "\n\n")
```
---

# 5. å“è³ªè©•ä¾¡ã¨æ¤œè¨¼ {.tabset .tabset-fade}

## ãƒžãƒ¼ã‚«ãƒ¼éºä¼å­ç™ºç¾ç¢ºèª

```{r marker_expression_validation}
cat("=== Validating Marker Gene Expression ===\n")

# Key markers to validate
key_markers <- list(
  PTPRC = "Immune",
  CD3D = "T_cell",
  CD19 = "B_cell",
  CD68 = "Macrophage",
  LRP2 = "Proximal_Tubule",
  SLC12A3 = "Distal_Tubule",
  AQP2 = "Collecting_Duct",
  NPHS2 = "Podocyte",
  PECAM1 = "Endothelial",
  COL1A1 = "Stromal"
)

# Check expression in predicted cell types
DefaultAssay(cosmx) <- "SCT"
validation_results <- data.frame()

for(marker in names(key_markers)) {
  expected_celltype <- key_markers[[marker]]
  
  if(marker %in% rownames(cosmx)) {
    expr <- GetAssayData(cosmx, assay = "SCT", layer = "data")[marker, ]
    
    # Expression in expected cell type
    if(grepl("_", expected_celltype)) {
      # Detailed cell type
      expected_cells <- cosmx$detailed_celltype == expected_celltype
    } else {
      # Major cell type or contains pattern
      expected_cells <- cosmx$major_celltype == expected_celltype | 
                       grepl(expected_celltype, cosmx$detailed_celltype)
    }
    
    if(sum(expected_cells) > 0) {
      expr_in_expected <- expr[expected_cells]
      expr_in_others <- expr[!expected_cells]
      
      validation_results <- rbind(validation_results, data.frame(
        Marker = marker,
        Expected_CellType = expected_celltype,
        N_Expected = sum(expected_cells),
        Mean_Expr_Expected = round(mean(expr_in_expected), 3),
        Pct_Pos_Expected = round(sum(expr_in_expected > 0) / length(expr_in_expected) * 100, 1),
        Mean_Expr_Others = round(mean(expr_in_others), 3),
        Pct_Pos_Others = round(sum(expr_in_others > 0) / length(expr_in_others) * 100, 1),
        Fold_Change = round(mean(expr_in_expected) / (mean(expr_in_others) + 0.001), 2)
      ))
    }
  }
}

print(validation_results)

# Save validation results
write.csv(validation_results,
          file.path(output_dir, "marker_validation_results.csv"),
          row.names = FALSE)

cat("\nâœ… Marker validation complete\n\n")
```

## ã‚µãƒ³ãƒ—ãƒ«åˆ¥åˆ†å¸ƒ

```{r sample_distribution}
cat("=== Sample-wise Cell Type Distribution ===\n")

# Major cell types by sample
major_by_sample <- table(cosmx$sample, cosmx$major_celltype)
cat("=== Major Cell Types by Sample ===\n")
print(major_by_sample)

# Percentage
major_by_sample_pct <- prop.table(major_by_sample, margin = 1) * 100
print(round(major_by_sample_pct, 2))

# Save
write.csv(major_by_sample,
          file.path(output_dir, "major_celltype_by_sample.csv"))

# Detailed cell types by sample (å®Ÿéš›ã®æ•°ã«å¯¾å¿œ)
detailed_by_sample <- table(cosmx$sample, cosmx$detailed_celltype)

# å®Ÿéš›ã®ç´°èƒžã‚¿ã‚¤ãƒ—æ•°ã‚’ç¢ºèª
n_celltypes <- ncol(detailed_by_sample)
top_n <- min(20, n_celltypes)  # 20ã¨å®Ÿéš›ã®æ•°ã®å°ã•ã„æ–¹

top_celltypes <- names(sort(table(cosmx$detailed_celltype), decreasing = TRUE)[1:top_n])
detailed_by_sample_top <- detailed_by_sample[, top_celltypes]

cat("\n=== Top", top_n, "Detailed Cell Types by Sample ===\n")
print(detailed_by_sample_top)

# Save
write.csv(detailed_by_sample_top,
          file.path(output_dir, paste0("top", top_n, "_detailed_celltype_by_sample.csv")))

cat("\n")
```

## ç©ºé–“åˆ†å¸ƒç¢ºèª

```{r spatial_distribution_check}
cat("=== Checking Spatial Distribution ===\n")

# Check for different coordinate systems
has_fov_coords <- all(c("x_FOV_px", "y_FOV_px", "fov") %in% colnames(cosmx@meta.data))
has_slide_coords <- all(c("x_slide_mm", "y_slide_mm") %in% colnames(cosmx@meta.data))

if(has_slide_coords) {
  
  cat("âœ… Slide coordinates available (x_slide_mm, y_slide_mm)\n")
  
  # Plot spatial distribution for each sample using slide coordinates
  for(sample_id in unique(cosmx$sample)) {
    sample_cells <- cosmx$sample == sample_id
    sample_data <- cosmx@meta.data[sample_cells, ]
    
    p <- ggplot(sample_data, aes(x = x_slide_mm, y = y_slide_mm, 
                                  color = major_celltype)) +
      geom_point(size = 0.1, alpha = 0.6) +
      scale_color_brewer(palette = "Set2") +
      theme_minimal() +
      theme(legend.position = "right",
            aspect.ratio = 1) +
      labs(title = paste("Sample", sample_id, "- Major Cell Types"),
           x = "X coordinate (mm)",
           y = "Y coordinate (mm)",
           color = "Cell Type")
    
    print(p)
    
    ggsave(
      file.path(output_dir, paste0("spatial_major_celltype_", sample_id, ".png")),
      p, width = 10, height = 8, dpi = 300
    )
  }
  
  cat("âœ… Spatial distribution plots saved\n")
  
} else if(has_fov_coords) {
  
  cat("âœ… FOV coordinates available (x_FOV_px, y_FOV_px)\n")
  cat("   Note: Plotting by FOV - multiple panels per sample\n")
  
  # Plot by FOV
  for(sample_id in unique(cosmx$sample)) {
    sample_cells <- cosmx$sample == sample_id
    sample_data <- cosmx@meta.data[sample_cells, ]
    
    p <- ggplot(sample_data, aes(x = x_FOV_px, y = y_FOV_px, 
                                  color = major_celltype)) +
      geom_point(size = 0.1, alpha = 0.6) +
      facet_wrap(~fov, ncol = 5) +
      scale_color_brewer(palette = "Set2") +
      theme_minimal() +
      theme(legend.position = "bottom",
            strip.text = element_text(size = 6)) +
      labs(title = paste("Sample", sample_id, "- Major Cell Types (by FOV)"),
           x = "X coordinate (pixels)",
           y = "Y coordinate (pixels)",
           color = "Cell Type")
    
    print(p)
    
    ggsave(
      file.path(output_dir, paste0("spatial_fov_major_celltype_", sample_id, ".png")),
      p, width = 16, height = 12, dpi = 300
    )
  }
  
  cat("âœ… Spatial distribution plots (FOV) saved\n")
  
} else {
  cat("âš ï¸ Spatial coordinates not found\n")
  cat("   Expected: x_global_px/y_global_px or x_slide_mm/y_slide_mm or x_FOV_px/y_FOV_px\n")
}

cat("\n")
```

---

# 6. å¯è¦–åŒ– {.tabset .tabset-fade}

## UMAP: Major Cell Types

```{r umap_major_celltypes}
cat("=== UMAP Visualization: Major Cell Types ===\n")

# Check if UMAP exists
if("umap.harmony.sct" %in% names(cosmx@reductions)) {
  
  p_umap_major <- DimPlot(
    cosmx,
    reduction = "umap.harmony.sct",
    group.by = "major_celltype",
    label = TRUE,
    label.size = 4,
    pt.size = 0.3,
    raster = FALSE
  ) +
    scale_color_brewer(palette = "Set2") +
    theme_minimal(base_size = 14) +
    labs(title = "Major Cell Types (UMAP)",
         x = "UMAP 1",
         y = "UMAP 2") +
    theme(plot.title = element_text(hjust = 0.5, face = "bold"))
  
  print(p_umap_major)
  
  ggsave(
    file.path(output_dir, "umap_major_celltypes.png"),
    p_umap_major, width = 10, height = 8, dpi = 300
  )
  
  cat("âœ… UMAP major cell types saved\n")
  
} else {
  cat("âš ï¸ UMAP reduction not found\n")
}

cat("\n")
```

## UMAP: Detailed Cell Types

```{r umap_detailed_celltypes}
cat("=== UMAP Visualization: Detailed Cell Types ===\n")

if("umap.harmony.sct" %in% names(cosmx@reductions)) {
  
  # Plot top 15 cell types only for clarity
  top15_types <- names(sort(table(cosmx$detailed_celltype), decreasing = TRUE)[1:15])
  cosmx$detailed_celltype_top15 <- ifelse(
    cosmx$detailed_celltype %in% top15_types,
    as.character(cosmx$detailed_celltype),
    "Other"
  )
  
  p_umap_detailed <- DimPlot(
    cosmx,
    reduction = "umap.harmony.sct",
    group.by = "detailed_celltype_top15",
    label = TRUE,
    label.size = 3,
    pt.size = 0.3,
    raster = FALSE
  ) +
    theme_minimal(base_size = 12) +
    labs(title = "Detailed Cell Types - Top 15 (UMAP)",
         x = "UMAP 1",
         y = "UMAP 2") +
    theme(plot.title = element_text(hjust = 0.5, face = "bold"),
          legend.position = "right")
  
  print(p_umap_detailed)
  
  ggsave(
    file.path(output_dir, "umap_detailed_celltypes_top15.png"),
    p_umap_detailed, width = 12, height = 8, dpi = 300
  )
  
  cat("âœ… UMAP detailed cell types saved\n")
  
}

cat("\n")
```

## Module Score Heatmap

```{r module_score_heatmap}
cat("=== Module Score Heatmap ===\n")

# Get score columns
score_cols <- grep("_Score$", colnames(cosmx@meta.data), value = TRUE)

if(length(score_cols) > 0) {
  
  # Sample cells for heatmap (max 5000 per major cell type)
  set.seed(123)
  sampled_cells <- c()
  for(ct in unique(cosmx$major_celltype)) {
    ct_cells <- which(cosmx$major_celltype == ct)
    n_sample <- min(5000, length(ct_cells))
    sampled_cells <- c(sampled_cells, sample(ct_cells, n_sample))
  }
  
  # Extract scores
  score_matrix <- t(as.matrix(cosmx@meta.data[sampled_cells, score_cols]))
  
  # Create unique column names using cell indices
  colnames(score_matrix) <- paste0("Cell_", sampled_cells)
  
  # Clean row names (remove _Score suffix)
  rownames(score_matrix) <- gsub("_Score$", "", rownames(score_matrix))
  
  # Create annotation data frame with unique row names
  annotation_df <- data.frame(
    Major_CellType = cosmx$major_celltype[sampled_cells],
    row.names = paste0("Cell_", sampled_cells)
  )
  
  # Define colors for cell types
  cell_type_colors <- c(
    "Epithelial" = "#66C2A5",
    "Other" = "#FC8D62",
    "Stromal" = "#8DA0CB",
    "Endothelial" = "#E78AC3",
    "Immune" = "#A6D854"
  )
  
  # Plot heatmap
  pdf(file.path(output_dir, "module_score_heatmap.pdf"), width = 12, height = 10)
  pheatmap(
    score_matrix,
    show_colnames = FALSE,
    cluster_rows = TRUE,
    cluster_cols = TRUE,
    color = colorRampPalette(c("blue", "white", "red"))(100),
    main = "Module Scores by Cell Type",
    fontsize_row = 8,
    annotation_col = annotation_df,
    annotation_colors = list(Major_CellType = cell_type_colors)
  )
  dev.off()
  
  cat("âœ… Module score heatmap saved\n")
}

cat("\n")
```

## ä¸»è¦ãƒžãƒ¼ã‚«ãƒ¼ç™ºç¾

```{r key_marker_expression}
cat("=== Key Marker Expression ===\n")

# Key markers to visualize
markers_to_plot <- c("PTPRC", "CD3D", "CD68", "LRP2", "SLC12A3", 
                     "AQP2", "NPHS2", "PECAM1", "COL1A1")
markers_available <- intersect(markers_to_plot, rownames(cosmx))

if(length(markers_available) > 0 && "umap.harmony.sct" %in% names(cosmx@reductions)) {
  
  # Create feature plots
  p_markers <- FeaturePlot(
    cosmx,
    features = markers_available,
    reduction = "umap.harmony.sct",
    ncol = 3,
    pt.size = 0.3,
    order = TRUE
  )
  
  print(p_markers)
  
  ggsave(
    file.path(output_dir, "key_marker_expression_umap.png"),
    p_markers, width = 15, height = ceiling(length(markers_available)/3) * 5, dpi = 300
  )
  
  cat("âœ… Key marker expression plots saved\n")
}

cat("\n")
```

---

# 7. ãƒ‡ãƒ¼ã‚¿ä¿å­˜ {.tabset .tabset-fade}

## ãƒ¡ã‚¿ãƒ‡ãƒ¼ã‚¿ä¿å­˜

```{r save_metadata}
cat("=== Saving Metadata ===\n")

# Check if condition column exists, if not create it
if(!"condition" %in% colnames(cosmx@meta.data)) {
  cat("âš ï¸  'condition' column not found - creating from sample info\n")
  cosmx@meta.data$condition <- case_when(
    cosmx@meta.data$sample == "S0" ~ "Rejection",
    cosmx@meta.data$sample == "S1" ~ "ICI-AKI",
    cosmx@meta.data$sample == "S2" ~ "Drug-induced_AKI",
    cosmx@meta.data$sample == "S3" ~ "ICI-AKI",
    TRUE ~ "Unknown"
  )
  cat("âœ… 'condition' column created\n")
}

# Check if spatial coordinates exist
spatial_cols <- c()
if(all(c("x_slide_mm", "y_slide_mm") %in% colnames(cosmx@meta.data))) {
  spatial_cols <- c("x_slide_mm", "y_slide_mm")
  cat("âœ… Using slide coordinates (mm)\n")
} else if(all(c("x_FOV_px", "y_FOV_px") %in% colnames(cosmx@meta.data))) {
  spatial_cols <- c("x_FOV_px", "y_FOV_px")
  cat("âœ… Using FOV coordinates (pixels)\n")
} else {
  cat("âš ï¸  No spatial coordinates found\n")
}

# Extract key metadata columns
metadata_export <- cosmx@meta.data %>%
  select(
    # Sample info
    sample, condition,
    # Cell type annotations
    major_celltype, immune_subtype, epithelial_subtype, detailed_celltype,
    # Spatial coordinates (if available)
    any_of(spatial_cols),
    # QC metrics
    nCount_SCT, nFeature_SCT,
    # All module scores
    ends_with("_Score")
  )

# Save as CSV
write.csv(
  metadata_export,
  file.path(output_dir, "cosmx_comprehensive_annotation_metadata.csv"),
  row.names = TRUE
)

cat("âœ… Metadata saved\n")
cat("Rows:", nrow(metadata_export), "\n")
cat("Columns:", ncol(metadata_export), "\n\n")
```

## Seuratã‚ªãƒ–ã‚¸ã‚§ã‚¯ãƒˆä¿å­˜

```{r save_seurat_object}
cat("=== Saving Annotated Seurat Object ===\n")

# Save complete annotated object
output_rds <- file.path(data_dir, "20251117_CosMx_SCT_nolabeltransfer_comprehensive_annotation.rds")

cat("Saving to:", output_rds, "\n")
cat("Object size:", format(object.size(cosmx), units = "GB"), "\n")

saveRDS(cosmx, output_rds)

cat("âœ… Seurat object saved\n")
cat("File size:", round(file.info(output_rds)$size / 1024^3, 2), "GB\n\n")
```

## ã‚µãƒžãƒªãƒ¼ãƒ¬ãƒãƒ¼ãƒˆ

```{r summary_report}
cat("=== Generating Summary Report ===\n")

# Create summary report
summary_report <- list(
  analysis_date = Sys.Date(),
  total_cells = ncol(cosmx),
  total_genes = nrow(cosmx),
  samples = unique(cosmx$sample),
  
  major_celltypes = list(
    count = length(unique(cosmx$major_celltype)),
    distribution = as.list(table(cosmx$major_celltype))
  ),
  
  detailed_celltypes = list(
    count = length(unique(cosmx$detailed_celltype)),
    top10 = names(sort(table(cosmx$detailed_celltype), decreasing = TRUE)[1:10])
  ),
  
  module_scores_calculated = length(grep("_Score$", colnames(cosmx@meta.data))),
  
  marker_genes_used = list(
    epithelial = length(unlist(epithelial_available)),
    immune = length(unlist(immune_available)),
    stromal = length(unlist(stromal_available))
  )
)

# Save as RDS
saveRDS(summary_report, file.path(output_dir, "analysis_summary_report.rds"))

# Print summary
cat("\n")
cat(strrep("=", 70), "\n")
cat("COMPREHENSIVE ANNOTATION SUMMARY\n")
cat(strrep("=", 70), "\n")
cat("Analysis Date:", as.character(summary_report$analysis_date), "\n")
cat("Total Cells:", summary_report$total_cells, "\n")
cat("Total Genes:", summary_report$total_genes, "\n")
cat("Samples:", paste(summary_report$samples, collapse = ", "), "\n")
cat("\nMajor Cell Types:", summary_report$major_celltypes$count, "\n")
for(ct in names(summary_report$major_celltypes$distribution)) {
  cat(sprintf("  %-15s: %6d (%.2f%%)\n", 
              ct, 
              summary_report$major_celltypes$distribution[[ct]],
              summary_report$major_celltypes$distribution[[ct]] / summary_report$total_cells * 100))
}
cat("\nDetailed Cell Types:", summary_report$detailed_celltypes$count, "\n")
cat("Top 10:\n")
for(i in 1:min(10, length(summary_report$detailed_celltypes$top10))) {
  ct <- summary_report$detailed_celltypes$top10[i]
  count <- sum(cosmx$detailed_celltype == ct)
  cat(sprintf("  %2d. %-25s: %6d (%.2f%%)\n", 
              i, ct, count, count / summary_report$total_cells * 100))
}
cat("\nModule Scores Calculated:", summary_report$module_scores_calculated, "\n")
cat(strrep("=", 70), "\n")

cat("\nâœ… Analysis complete!\n")
cat("Output directory:", output_dir, "\n\n")
```

---

# 8. ã‚»ãƒƒã‚·ãƒ§ãƒ³æƒ…å ±

```{r session_info}
cat("=== Session Information ===\n")
sessionInfo()
```

---

# ã¾ã¨ã‚

## ä¸»è¦ãªæˆæžœ

1. **CosMx 1,000éºä¼å­ãƒ‘ãƒãƒ«ã®å®Œå…¨æ´»ç”¨**
   - ä¸Šçš®ç´°èƒžãƒžãƒ¼ã‚«ãƒ¼: 7ã‚«ãƒ†ã‚´ãƒª
   - å…ç–«ç´°èƒžãƒžãƒ¼ã‚«ãƒ¼: 11ã‚«ãƒ†ã‚´ãƒª
   - é–“è³ªãƒ»è¡€ç®¡ç³»ãƒžãƒ¼ã‚«ãƒ¼: 6ã‚«ãƒ†ã‚´ãƒª

2. **éšŽå±¤çš„ç´°èƒžã‚¿ã‚¤ãƒ—åˆ†é¡ž**
   - Level 1: Major cell types (5-8ã‚¿ã‚¤ãƒ—)
   - Level 2: Detailed cell types (20-30ã‚¿ã‚¤ãƒ—)

3. **å“è³ªæ¤œè¨¼**
   - ãƒžãƒ¼ã‚«ãƒ¼éºä¼å­ç™ºç¾ã®ä¸€è²«æ€§ç¢ºèª
   - ã‚µãƒ³ãƒ—ãƒ«é–“åˆ†å¸ƒã®å¦¥å½“æ€§ç¢ºèª
   - ç©ºé–“åˆ†å¸ƒãƒ‘ã‚¿ãƒ¼ãƒ³ã®æ¤œè¨¼

4. **åŒ…æ‹¬çš„ãƒ‡ãƒ¼ã‚¿ã‚»ãƒƒãƒˆä½œæˆ**
   - å…¨ç´°èƒžã«å¯¾ã™ã‚‹è©³ç´°ã‚¢ãƒŽãƒ†ãƒ¼ã‚·ãƒ§ãƒ³
   - Module scoreå®Œå‚™
   - ç©ºé–“åº§æ¨™æƒ…å ±ä¿æŒ

## æ¬¡ã®ã‚¹ãƒ†ãƒƒãƒ—

1. **Immuneç´°èƒžã®è©³ç´°è§£æž**
   - CD45+ç´°èƒžã®å†ã‚¯ãƒ©ã‚¹ã‚¿ãƒªãƒ³ã‚°
   - Tç´°èƒžã‚µãƒ–ã‚»ãƒƒãƒˆåˆ†é¡ž
   - ICI-AKIç‰¹ç•°çš„ãƒžãƒ¼ã‚«ãƒ¼æŽ¢ç´¢

2. **ä¸Šçš®ç´°èƒžã®æ¤œè¨¼**
   - å„ã‚µãƒ–ã‚¿ã‚¤ãƒ—ã®ç©ºé–“åˆ†å¸ƒç¢ºèª
   - å°¿ç´°ç®¡ã‚»ã‚°ãƒ¡ãƒ³ãƒˆåŒå®š
   - ç—…æ…‹é–¢é€£éºä¼å­ç™ºç¾è§£æž

3. **çµ±åˆè§£æž**
   - scFFPEãƒ‡ãƒ¼ã‚¿ã¨ã®æ¯”è¼ƒæ¤œè¨¼
   - ãƒã‚¤ã‚ªãƒžãƒ¼ã‚«ãƒ¼åŒå®š
   - ç—…æ…‹ãƒ¡ã‚«ãƒ‹ã‚ºãƒ è§£æ˜Ž

---

**Analysis completed**: `r Sys.time()`

