---
title: "scFFPE Figure 2 & CosMx Figure 4: ICI-AKI Signature Discovery and Spatial Coupling"
author: "Kidney ICI-AKI Research Project"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
    code_folding: show
    theme: cosmo
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 12,
  fig.height = 8,
  dpi = 300
)
```

# ãƒ—ãƒ­ã‚¸ã‚§ã‚¯ãƒˆæ¦‚è¦

## ç›®çš„
1. **scFFPE Figure 2**: ICIç‰¹ç•°çš„ã‚·ã‚°ãƒãƒãƒ£ãƒ¼ã®ã‚½ãƒ¼ã‚¹ã‚¨ãƒ“ãƒ‡ãƒ³ã‚¹ã‚’ç¤ºã™
   - 2A: Workflow schematicï¼ˆæ‰‹å‹•ä½œæˆæ¸ˆã¿ï¼‰
   - 2B: scFFPE UMAP (ä¸»è¦ç´°èƒžã‚¿ã‚¤ãƒ—ã¨ã‚·ã‚°ãƒãƒãƒ£ãƒ¼éºä¼å­)
   - 2C: scFFPE DEG heatmap
   
2. **CosMx Figure 4**: Tç´°èƒžÃ—éª¨é«„ç³»ç´°èƒžã®ç©ºé–“çš„ç›¸äº’ä½œç”¨
   - 4A: Nearest-neighbor Tâ†’Myeloid distance
   - 4B: Co-localization index
   - 4C: Exhaustion-high T cells Ã— Myeloid-Activation-high overlap
   - 4D: Spatial correlation (Ripley's K or radial density)

## ãƒ‡ãƒ¼ã‚¿æ§‹é€ 
- **scFFPE**: `/Users/HiroyukiArai/Documents/å¤§å­¦é™¢/ç ”ç©¶çµæžœ/2024æ‘ä¸Šå…ˆç”ŸCosMX/202406 single cell FFPE data/scFFPE_kidney_analysis/results/scffpe_additional_analysis_complete_doublet_removed.rds`
- **CosMx**: `/Users/HiroyukiArai/Documents/å¤§å­¦é™¢/ç ”ç©¶çµæžœ/2024æ‘ä¸Šå…ˆç”ŸCosMX/20251117_CosMx_SCT_nolabeltransfer_comprehensive_annotation.rds`
- **Signatures**: `ICI_signature_candidates_for_CosMx.rds`

---

# 1. ç’°å¢ƒæº–å‚™ {.tabset .tabset-fade}

## ãƒ©ã‚¤ãƒ–ãƒ©ãƒªèª­ã¿è¾¼ã¿

```{r load_libraries}
cat("=== Loading Required Libraries ===\n")

suppressPackageStartupMessages({
  # Core
  library(Seurat)
  library(SeuratObject)
  library(dplyr)
  library(tidyr)
  library(tibble)
  
  # Visualization
  library(ggplot2)
  library(patchwork)
  library(viridis)
  library(RColorBrewer)
  library(pheatmap)
  library(ggrepel)
  
  # Spatial analysis
  library(spatstat)
  library(dbscan)
  
  # Statistics
  library(ggpubr)
  library(rstatix)
  
  # Data handling
  library(readxl)
  library(DT)
  library(kableExtra)
})

cat("âœ… All libraries loaded\n")
cat("Seurat version:", as.character(packageVersion("Seurat")), "\n")
cat("R version:", R.version.string, "\n\n")
```

## ãƒ‡ã‚£ãƒ¬ã‚¯ãƒˆãƒªè¨­å®š

```{r setup_directories}
cat("=== Setting up Directories ===\n")

# 2ã¤ã®å¯èƒ½æ€§ã®ã‚ã‚‹ãƒ™ãƒ¼ã‚¹ãƒ‡ã‚£ãƒ¬ã‚¯ãƒˆãƒªã‚’è©¦è¡Œ
base_dir_options <- c(
  "/Users/HiroyukiArai/Documents/2025å¹´ç•™å­¦/ç ”ç©¶é–¢é€£/ç ”ç©¶çµæžœ/202509 cosmx-kidney-project",
  "/Users/HiroyukiArai/Documents/å¤§å­¦é™¢/ç ”ç©¶çµæžœ/2024æ‘ä¸Šå…ˆç”ŸCosMX"
)

base_dir <- NULL
for (dir in base_dir_options) {
  if (dir.exists(dir)) {
    base_dir <- dir
    break
  }
}

if (is.null(base_dir)) {
  stop("âŒ Base directory not found. Please check the path.")
}

cat("âœ… Using base directory:", base_dir, "\n")

# scFFPE directory
scffpe_base <- "/Users/HiroyukiArai/Documents/å¤§å­¦é™¢/ç ”ç©¶çµæžœ/2024æ‘ä¸Šå…ˆç”ŸCosMX/202406 single cell FFPE data/scFFPE_kidney_analysis"
scffpe_data_dir <- file.path(scffpe_base, "results")
scffpe_deg_dir <- file.path(scffpe_base, "results/DEG_analysis")

# CosMx directory
cosmx_data_dir <- base_dir

# Output directory - timestamped
timestamp <- format(Sys.time(), "%Y%m%d_%H%M%S")
output_dir <- file.path(base_dir, "results/figures", 
                        paste0("20251203_Figure2(scFFPE_heatmap)", timestamp))
dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)

# Create subdirectories
dir.create(file.path(output_dir, "Figure2"), showWarnings = FALSE)
dir.create(file.path(output_dir, "Statistics"), showWarnings = FALSE)
dir.create(file.path(output_dir, "Additional"), showWarnings = FALSE)
cat("\nDirectory structure:\n")
cat("  scFFPE data:", scffpe_data_dir, "\n")
cat("  scFFPE DEG:", scffpe_deg_dir, "\n")
cat("  CosMx data:", cosmx_data_dir, "\n")
cat("  Output:", output_dir, "\n\n")
```

## ã‚«ãƒ©ãƒ¼ãƒ‘ãƒ¬ãƒƒãƒˆå®šç¾©

```{r define_colors}
cat("=== Defining Color Palettes ===\n")

# Condition colors
condition_colors <- c(
  "ICI-AKI" = "#E41A1C",
  "ICI_AKI" = "#E41A1C",
  "Rejection" = "#377EB8",
  "Drug-induced_AKI" = "#4DAF4A",
  "Drug_AKI" = "#4DAF4A"
)

# Sample colors
sample_colors <- c(
  "S0" = "#377EB8",
  "S1" = "#E41A1C",
  "S2" = "#4DAF4A",
  "S3" = "#984EA3"
)

# Major cell type colors
major_celltype_colors <- c(
  "Epithelial" = "#E41A1C",
  "Tubular_cells" = "#E41A1C",
  "Endothelial" = "#377EB8",
  "Immune" = "#4DAF4A",
  "Immune_cells" = "#4DAF4A",
  "Stromal" = "#984EA3",
  "Stromal_cells" = "#984EA3",
  "Other" = "#999999"
)

# Immune cell type colors
immune_colors <- c(
  "CD4_T_cell" = "#E41A1C",
  "CD8_T_cell" = "#377EB8",
  "DP_T_cell" = "#4DAF4A",
  "T_cell" = "#984EA3",
  "Treg" = "#FF7F00",
  "B_cell" = "#FFFF33",
  "Macrophage" = "#A65628",
  "Dendritic_cell" = "#F781BF",
  "NK_cell" = "#999999"
)

cat("âœ… Color palettes defined\n\n")
```

## è«–æ–‡ç”¨ãƒ†ãƒ¼ãƒž

```{r plot_theme}
cat("=== Defining Publication Theme ===\n")

# Base theme for all plots
theme_publication <- function(base_size = 12) {
  theme_bw(base_size = base_size) +
    theme(
      # Text
      plot.title = element_text(size = base_size + 2, face = "bold", hjust = 0.5),
      plot.subtitle = element_text(size = base_size, hjust = 0.5),
      axis.title = element_text(size = base_size, face = "bold"),
      axis.text = element_text(size = base_size - 1, color = "black"),
      
      # Legend
      legend.title = element_text(size = base_size, face = "bold"),
      legend.text = element_text(size = base_size - 1),
      legend.background = element_rect(fill = "white", color = "black"),
      legend.key.size = unit(0.5, "cm"),
      
      # Panel
      panel.background = element_rect(fill = "white"),
      panel.grid.major = element_line(color = "grey90", size = 0.3),
      panel.grid.minor = element_blank(),
      panel.border = element_rect(color = "black", fill = NA, size = 1),
      
      # Strip (facet labels)
      strip.background = element_rect(fill = "grey90", color = "black"),
      strip.text = element_text(size = base_size, face = "bold"),
      
      # Margins
      plot.margin = margin(10, 10, 10, 10)
    )
}

cat("âœ… Publication theme defined\n\n")
```

---

# 2. ãƒ‡ãƒ¼ã‚¿èª­ã¿è¾¼ã¿ {.tabset .tabset-fade}

## scFFPEãƒ‡ãƒ¼ã‚¿èª­ã¿è¾¼ã¿

```{r load_scffpe}
cat("=== Loading scFFPE Data ===\n")

scffpe_file <- file.path(scffpe_data_dir, "scffpe_additional_analysis_complete_doublet_removed.rds")

if (!file.exists(scffpe_file)) {
  stop("âŒ scFFPE file not found: ", scffpe_file)
}

cat("Loading scFFPE object...\n")
scffpe <- readRDS(scffpe_file)

cat("âœ… scFFPE data loaded\n")
cat("Total cells:", ncol(scffpe), "\n")
cat("Total genes:", nrow(scffpe), "\n")
cat("Assays:", names(scffpe@assays), "\n")

# Check metadata
cat("\nAvailable metadata columns:\n")
print(colnames(scffpe@meta.data))

# Create group column if not exists
if (!"group" %in% colnames(scffpe@meta.data)) {
  cat("\nCreating 'group' column...\n")
  
  # Check sample column name
  sample_col <- if ("sample" %in% colnames(scffpe@meta.data)) {
    "sample"
  } else if ("orig.ident" %in% colnames(scffpe@meta.data)) {
    "orig.ident"
  } else {
    stop("Cannot find sample information column")
  }
  
  # Create group
  scffpe$group <- case_when(
    grepl("S0", scffpe@meta.data[[sample_col]]) ~ "Rejection",
    grepl("S1|S3", scffpe@meta.data[[sample_col]]) ~ "ICI_AKI",
    grepl("S2", scffpe@meta.data[[sample_col]]) ~ "Drug_AKI",
    TRUE ~ "Unknown"
  )
  
  cat("âœ… 'group' column created\n")
  print(table(scffpe$group))
}

cat("\n")
```

## CosMxãƒ‡ãƒ¼ã‚¿èª­ã¿è¾¼ã¿
```{r load_cosmx}
cat("=== Loading CosMx Data ===\n")

# ã‚·ã‚°ãƒãƒãƒ£ãƒ¼ã‚¹ã‚³ã‚¢ä»˜ãCosMxãƒ•ã‚¡ã‚¤ãƒ«ï¼ˆFigure 3ã¨åŒã˜ï¼‰
cosmx_file <- "/Users/HiroyukiArai/Documents/2025å¹´ç•™å­¦/ç ”ç©¶é–¢é€£/ç ”ç©¶çµæžœ/202509 cosmx-kidney-project/results/figures/20251120_CosMx_Signature_Projection_20251120_175233/CosMx_with_signatures_20251120_175944.rds"

if (!file.exists(cosmx_file)) {
  stop("âŒ CosMx file not found: ", cosmx_file)
}

cat("Loading CosMx object from:", basename(cosmx_file), "\n")
cosmx <- readRDS(cosmx_file)

cat("âœ… CosMx data loaded\n")
cat("Total cells:", ncol(cosmx), "\n")
cat("Total genes:", nrow(cosmx), "\n")
cat("Samples:", paste(unique(cosmx$sample), collapse = ", "), "\n")

# Check metadata
cat("\nAvailable metadata columns:\n")
print(colnames(cosmx@meta.data))

# Check for signature scores
sig_cols <- grep("_1$", colnames(cosmx@meta.data), value = TRUE)
cat("\n=== Signature Scores Available ===\n")
print(sig_cols)

# Check for spatial coordinates
if ("x_slide_mm" %in% colnames(cosmx@meta.data)) {
  cat("\nâœ… Using x_slide_mm, y_slide_mm (global slide coordinates)\n")
  x_col <- "x_slide_mm"
  y_col <- "y_slide_mm"
} else if ("x_global_px" %in% colnames(cosmx@meta.data)) {
  cat("âœ… Using x_global_px, y_global_px\n")
  x_col <- "x_global_px"
  y_col <- "y_global_px"
} else if ("x_FOV_px" %in% colnames(cosmx@meta.data)) {
  cat("âš ï¸ Using x_FOV_px, y_FOV_px (FOV-level coordinates)\n")
  x_col <- "x_FOV_px"
  y_col <- "y_FOV_px"
} else {
  stop("âŒ No coordinate columns found")
}

cat("\n")
```

## Signatureèª­ã¿è¾¼ã¿

```{r load_signatures}
cat("=== Loading ICI Signatures ===\n")

signature_file <- file.path(scffpe_deg_dir, "ICI_signature_candidates_for_CosMx.rds")

if (!file.exists(signature_file)) {
  stop("âŒ Signature file not found: ", signature_file)
}

signatures <- readRDS(signature_file)

cat("âœ… Signatures loaded\n")
cat("Number of signatures:", length(signatures), "\n")
cat("\nSignature names:\n")
for (i in seq_along(signatures)) {
  sig_name <- names(signatures)[i]
  sig_genes <- signatures[[i]]
  cat(sprintf("  %2d. %-30s: %2d genes\n", i, sig_name, length(sig_genes)))
}

cat("\n")
```

## DEGçµæžœèª­ã¿è¾¼ã¿

```{r load_deg_results}
cat("=== Loading DEG Results ===\n")

# ICI-specific genes
ici_specific_file <- file.path(scffpe_deg_dir, "ICI_specific_genes_all_cells.csv")

if (file.exists(ici_specific_file)) {
  ici_specific_genes <- read.csv(ici_specific_file)
  cat("âœ… ICI-specific genes loaded:", nrow(ici_specific_genes), "genes\n")
} else {
  cat("âš ï¸ ICI-specific genes file not found\n")
  ici_specific_genes <- NULL
}

# Cell type-specific DEGs
celltype_deg_file <- file.path(scffpe_deg_dir, "All_celltype_specific_DEGs.csv")

if (file.exists(celltype_deg_file)) {
  celltype_degs <- read.csv(celltype_deg_file)
  cat("âœ… Cell type-specific DEGs loaded:", nrow(celltype_degs), "entries\n")
} else {
  cat("âš ï¸ Cell type-specific DEG file not found\n")
  celltype_degs <- NULL
}

cat("\n")
```

---

# 3. Figure 2: scFFPE Signature Source Evidence {.tabset .tabset-fade}



## Figure 2C: ICI Signature Expression by Cell Type (Cell Type-grouped)
```{r figure2c_signature_heatmap_celltype_grouped, fig.width=12, fig.height=8}
cat("=== Creating Figure 2C: Cell Type-grouped Heatmap ===\n")

if (!is.null(ici_specific_genes)) {

## ---- STEP 1: Gene selection (å®Ÿéš›ã®Signatureã«åŸºã¥ã) ----

cat("\n=== Gene Selection Based on Actual CosMx Signatures ===\n")

# CosMxãƒ‘ãƒãƒ«éºä¼å­ã®å®šç¾©
if (exists("signatures") && !is.null(signatures)) {
  cosmx_panel_genes <- unique(unlist(signatures))
  cat("CosMx panel genes available:", length(cosmx_panel_genes), "\n")
} else {
  cosmx_panel_genes <- rownames(scffpe)
}

# å®Ÿéš›ã«ä½¿ç”¨ã—ãŸSignatureå®šç¾©ï¼ˆCosMxæ­è¼‰éºä¼å­ã®ã¿ï¼‰
gene_categories <- list(
  # T cell Exhaustionï¼ˆå®Ÿéš›ã®Signatureï¼‰
  "T_cell_Exhaustion" = c("ENTPD1", "CTLA4", "TIGIT"),
  
  # Inflammatory Chemokineï¼ˆå®Ÿéš›ã®Signatureï¼‰
  "Inflammatory_Chemokine" = c("CXCL12", "CXCL14", "IL17RB"),
  
  # Myeloid Activationï¼ˆå®Ÿéš›ã®Signatureï¼‰
  "Myeloid_Activation" = c("S100A9", "GPX3", "CD36", "VCAN", "G0S2", 
                           "CXCL14", "LTF", "SPP1", "FOS", "VEGFA"),
  
  # T cell Activationï¼ˆå®Ÿéš›ã®Signatureï¼‰
  "T_cell_Activation" = c("GPX3", "SPP1", "ENTPD1", "CRYAB", "TNFRSF4", 
                          "SERPINA1", "CTLA4", "TNFRSF18", "LTB", "EZH2"),
  
  # Tubular Injuryï¼ˆå®Ÿéš›ã®Signatureï¼‰
  "Tubular_Injury" = c("EGF", "GPX3", "PROX1", "G0S2", "MT1X", 
                       "DDC", "AZGP1", "APOE", "VEGFA", "LTF"),
  
  # Endothelial Activationï¼ˆå®Ÿéš›ã®Signatureï¼‰
  "Endothelial_Activation" = c("GPX3", "MT1X", "FOXF1", "FGFR3", "IGFBP5",
                                "SNAI2", "CEACAM1", "MRC2", "CXCL14", "SPP1"),
  
  # Stromal Activationï¼ˆå®Ÿéš›ã®Signatureï¼‰
  "Stromal_Activation" = c("GPX3", "CXCL14", "BMP5", "SPP1", "LDLR",
                           "CCL21", "G0S2", "TNFSF15", "VCAN", "FZD8"),
  
  # B/Plasma Activationï¼ˆå®Ÿéš›ã®Signatureï¼‰
  "B_Plasma_Activation" = c("GPX3", "MT1X", "CXCL14", "SPP1", "CRYAB",
                            "ADGRG1", "APOE", "VEGFA"),
  
  # Tissue Damageï¼ˆå®Ÿéš›ã®Signatureï¼‰
  "Tissue_Damage" = c("CXCL14", "SPP1", "VEGFA", "MT1X")
)

cat("\nSignature categories defined:\n")
for (cat_name in names(gene_categories)) {
  cat("  ", cat_name, ":", length(gene_categories[[cat_name]]), "genes\n")
}

# ICIç‰¹ç•°çš„ & scFFPEã§åˆ©ç”¨å¯èƒ½ & æ­£ã®log2FC
ici_available_genes <- ici_specific_genes %>%
  dplyr::filter(gene %in% rownames(scffpe)) %>%
  dplyr::filter(mean_log2FC > 0)

cat("\nICI-specific genes with positive log2FC:", nrow(ici_available_genes), "\n")

# å„ã‚«ãƒ†ã‚´ãƒªã‹ã‚‰éºä¼å­ã‚’é¸æŠž
selected_genes <- list()

for (cat_name in names(gene_categories)) {
  cat_genes <- gene_categories[[cat_name]]
  
  # ã“ã®ã‚«ãƒ†ã‚´ãƒªã®éºä¼å­ã§ã€ICIç‰¹ç•°çš„ & æ­£ã®log2FC & scFFPEã§åˆ©ç”¨å¯èƒ½
  available <- ici_available_genes %>%
    dplyr::filter(gene %in% cat_genes) %>%
    dplyr::arrange(dplyr::desc(mean_log2FC))
  
  if (nrow(available) > 0) {
    # å…¨éºä¼å­ã‚’å«ã‚ã‚‹ï¼ˆlog2FCé †ï¼‰
    selected_genes[[cat_name]] <- available$gene
    
    cat("  ", cat_name, ":", length(selected_genes[[cat_name]]), "genes selected\n")
  }
}

# çµ±åˆï¼ˆé‡è¤‡å‰Šé™¤ï¼‰
key_genes_available <- unique(unlist(selected_genes))

cat("\n=== Final Gene Set ===\n")
cat("Total unique genes:", length(key_genes_available), "\n")
cat("Genes by category:\n")
for (cat_name in names(selected_genes)) {
  if (length(selected_genes[[cat_name]]) > 0) {
    cat("  ", cat_name, ":\n    ", 
        paste(selected_genes[[cat_name]], collapse = ", "), "\n")
  }
}
cat("\n")

# 20-25éºä¼å­ã«èª¿æ•´ï¼ˆå¿…è¦ãªå ´åˆï¼‰
if (length(key_genes_available) > 25) {
  cat("Reducing to 25 genes...\n")
  
  # å„ªå…ˆé †ä½ï¼šT_cell_Exhaustion > Myeloid_Activation > Tubular_Injury
  priority_categories <- c("T_cell_Exhaustion", "Inflammatory_Chemokine",
                          "Myeloid_Activation", "Tubular_Injury", 
                          "T_cell_Activation", "Endothelial_Activation")
  
  key_genes_available <- c()
  for (cat_name in priority_categories) {
    if (cat_name %in% names(selected_genes)) {
      genes_to_add <- selected_genes[[cat_name]][1:min(4, length(selected_genes[[cat_name]]))]
      key_genes_available <- c(key_genes_available, genes_to_add)
      if (length(unique(key_genes_available)) >= 25) break
    }
  }
  key_genes_available <- unique(key_genes_available)[1:25]
}

cat("\nFinal gene count:", length(key_genes_available), "\n\n")

## ---- STEP 2: Broad Cell Type Ã— Condition pseudobulk ----
cat("\n=== Creating Pseudobulk Matrix by Broad Cell Type ===\n")

# major_celltypeã‚’çµ±ä¸€
scffpe$major_celltype_unified <- gsub("_cells$", "", scffpe$major_celltype)

# â˜…â˜…â˜… Broad_CellTypeã‚’äº‹å‰ã«å®šç¾© â˜…â˜…â˜…
scffpe$broad_celltype <- dplyr::case_when(
  grepl("Tubul", scffpe$major_celltype_unified, ignore.case = TRUE) ~ "Tubular",
  grepl("Myeloid", scffpe$major_celltype_unified, ignore.case = TRUE) ~ "Myeloid",
  grepl("T_NK|B_Plasma", scffpe$major_celltype_unified, ignore.case = TRUE) ~ "Immune",
  grepl("Endothel", scffpe$major_celltype_unified, ignore.case = TRUE) ~ "Endothelial",
  grepl("Glomerular", scffpe$major_celltype_unified, ignore.case = TRUE) ~ "Glomerular",
  grepl("Stromal", scffpe$major_celltype_unified, ignore.case = TRUE) ~ "Stromal",
  TRUE ~ "Other"
)

cat("Broad cell type distribution:\n")
print(table(scffpe$broad_celltype))
cat("\n")

## â˜… Broad_CellType Ã— Condition ã§pseudob ulkä½œæˆ
pseudobulk_list <- list()

for (broad_ct in unique(scffpe$broad_celltype)) {
  for (condition in unique(scffpe$group)) {

    # â˜… Broad_CellTypeã§ãƒ•ã‚£ãƒ«ã‚¿ï¼ˆT_NKã¨B_PlasmaãŒçµ±åˆã•ã‚Œã‚‹ï¼‰
    cells <- rownames(scffpe@meta.data)[
      scffpe@meta.data$broad_celltype == broad_ct &
        scffpe@meta.data$group == condition
    ]

    if (length(cells) >= 10) {
      expr_data <- GetAssayData(scffpe,
                                slot = "data",
                                assay = "RNA")[key_genes_available, cells,
                                               drop = FALSE]
      mean_expr <- rowMeans(expr_data)
      
      # â˜… ã‚µãƒ³ãƒ—ãƒ«åï¼šBroad_CellType_Condition
      sample_id <- paste0(broad_ct, "_", condition)
      pseudobulk_list[[sample_id]] <- mean_expr
      
      cat("Created:", sample_id, "with", length(cells), "cells\n")
    }
  }
}

if (length(pseudobulk_list) == 0) {
  cat("âš ï¸ No pseudobulk groups with â‰¥10 cells; skip Figure 2C.\n\n")
} else {

  pseudobulk_matrix <- do.call(cbind, pseudobulk_list)
  cat("\nPseudobulk matrix:", nrow(pseudobulk_matrix), "genes x",
      ncol(pseudobulk_matrix), "samples\n")
  
  cat("\nSample names:\n")
  print(colnames(pseudobulk_matrix))
  cat("\n")
  
## ---- STEP 3: Scaling & annotations ----

# gene-wise Z-score
pseudobulk_scaled <- t(scale(t(pseudobulk_matrix)))
pseudobulk_scaled[is.na(pseudobulk_scaled)] <- 0
pseudobulk_scaled[pseudobulk_scaled >  2.5] <-  2.5
pseudobulk_scaled[pseudobulk_scaled < -2.5] <- -2.5

# â˜… åˆ—ã‚¢ãƒŽãƒ†ãƒ¼ã‚·ãƒ§ãƒ³ï¼ˆã‚·ãƒ³ãƒ—ãƒ«ï¼‰
col_anno_df <- data.frame(
  Sample = colnames(pseudobulk_scaled),
  stringsAsFactors = FALSE
) %>%
  dplyr::mutate(
    # Broad_CellType: ã‚¢ãƒ³ãƒ€ãƒ¼ã‚¹ã‚³ã‚¢ã®å‰
    Broad_CellType = sub("_(Drug_AKI|ICI_AKI|Rejection)$", "", Sample),
    
    # Condition: ã‚¢ãƒ³ãƒ€ãƒ¼ã‚¹ã‚³ã‚¢ã®å¾Œ
    Condition = sub(".*_(Drug_AKI|ICI_AKI|Rejection)$", "\\1", Sample)
  )

rownames(col_anno_df) <- col_anno_df$Sample
col_anno_df$Sample <- NULL

cat("Column annotations:\n")
print(col_anno_df)
cat("\n")

# NAãƒã‚§ãƒƒã‚¯
if (any(is.na(col_anno_df$Broad_CellType)) || any(is.na(col_anno_df$Condition))) {
  stop("Sample name parsing failed")
}

## Condition colors
condition_colors_full <- c(
  "Drug_AKI"  = "#4DAF4A",
  "ICI_AKI"   = "#E41A1C",
  "Rejection" = "#377EB8"
)

present_conditions <- c("Drug_AKI", "ICI_AKI", "Rejection")
cond_cols <- condition_colors_full[present_conditions]

## Broad cell type colorsï¼ˆå›ºå®šé †åºï¼‰
broad_order <- c("Immune", "Myeloid", "Tubular", "Glomerular", "Endothelial", "Stromal", "Other")
present_broad <- intersect(broad_order, unique(col_anno_df$Broad_CellType))

fixed_colors <- c(
  "Immune"      = "#E78AC3",
  "Myeloid"     = "#FC8D62",
  "Tubular"     = "#FFD92F",
  "Glomerular"  = "#A6D854",
  "Endothelial" = "#66C2A5",
  "Stromal"     = "#8DA0CB",
  "Other"       = "#E5C494"
)

celltype_cols <- fixed_colors[present_broad]

## annotation_colors
ann_colors <- list(
  Cell_Type   = celltype_cols,
  Condition   = cond_cols
)

## ä¸¦ã³æ›¿ãˆï¼šBroad_CellType â†’ Condition
col_anno_df$Condition      <- factor(col_anno_df$Condition,
                                     levels = present_conditions)
col_anno_df$Broad_CellType <- factor(col_anno_df$Broad_CellType,
                                     levels = present_broad)

col_anno_df <- col_anno_df %>%
  dplyr::arrange(Broad_CellType, Condition)

cat("Final column order:\n")
print(col_anno_df)
cat("\n")

pseudobulk_scaled <- pseudobulk_scaled[, rownames(col_anno_df)]

# Cell Type ãƒ–ãƒ­ãƒƒã‚¯ã®åŒºåˆ‡ã‚Š
gap_positions <- which(diff(as.numeric(col_anno_df$Broad_CellType)) != 0)
cat("Gap positions:", gap_positions, "\n\n")

# xè»¸ãƒ©ãƒ™ãƒ«ï¼šConditionå
labels_col <- as.character(col_anno_df$Condition)

# pheatmap ç”¨ annotation_col
anno_col_for_pheatmap <- data.frame(
  Cell_Type = col_anno_df$Broad_CellType,
  Condition = col_anno_df$Condition,
  row.names = rownames(col_anno_df),
  stringsAsFactors = FALSE
)

## ---- STEP 4: Gene ordering and Heatmap (duplicate-safe) ----

# ã‚«ãƒ†ã‚´ãƒªåˆ¥ã«éºä¼å­ã‚’ã‚½ãƒ¼ãƒˆï¼ˆlog2FCé™é †ï¼‰
sorted_genes <- list()

for (cat_name in names(selected_genes)) {
  genes_in_cat <- intersect(selected_genes[[cat_name]], key_genes_available)
  
  if (length(genes_in_cat) > 0) {
    gene_fc <- ici_specific_genes %>%
      dplyr::filter(gene %in% genes_in_cat) %>%
      dplyr::arrange(dplyr::desc(mean_log2FC))
    
    sorted_genes[[cat_name]] <- gene_fc$gene
  }
}

cat("\n=== Gene ordering by category + log2FC (before dedup) ===\n")
for (cat_name in names(sorted_genes)) {
  if (length(sorted_genes[[cat_name]]) > 0) {
    cat(cat_name, ":", paste(sorted_genes[[cat_name]], collapse = ", "), "\n")
  }
}

# ---- ã“ã“ã‹ã‚‰é‡è¤‡è§£æ¶ˆãƒ­ã‚¸ãƒƒã‚¯ ----

# gene ã‚’ã©ã®ã‚«ãƒ†ã‚´ãƒªã«å‰²ã‚Šå½“ã¦ã‚‹ã‹ã®å„ªå…ˆé †ä½
priority_categories <- c(
  "T_cell_Exhaustion",
  "Inflammatory_Chemokine",
  "Myeloid_Activation",
  "T_cell_Activation",
  "Tubular_Injury",
  "Endothelial_Activation",
  "Stromal_Activation",
  "B_Plasma_Activation",
  "Tissue_Damage"
)

# ãƒ¦ãƒ‹ãƒ¼ã‚¯ãªéºä¼å­ã®é †ç•ªã¨ã€ãã®ã‚«ãƒ†ã‚´ãƒªã‚’æ±ºã‚ã‚‹
unique_genes <- c()
gene_category_vec <- c()

for (cat_name in priority_categories) {
  if (!cat_name %in% names(sorted_genes)) next
  
  for (g in sorted_genes[[cat_name]]) {
    if (!(g %in% unique_genes)) {
      unique_genes <- c(unique_genes, g)
      gene_category_vec[g] <- cat_name
    }
  }
}

gene_order <- unique_genes

cat("\n=== Final gene order (after dedup) ===\n")
print(gene_order)

# è¡Œã‚¢ãƒŽãƒ†ãƒ¼ã‚·ãƒ§ãƒ³ï¼ˆå„ gene ã« 1ã¤ã®ã‚«ãƒ†ã‚´ãƒªãƒ¼ï¼‰
gene_anno_df <- data.frame(
  Category = gene_category_vec[gene_order],
  row.names = gene_order,
  stringsAsFactors = FALSE
)

# ã‚«ãƒ†ã‚´ãƒªã”ã¨ã®é€£ç¶šé•·ã‹ã‚‰ã€è¡Œã®ã‚®ãƒ£ãƒƒãƒ—ä½ç½®ã‚’è¨ˆç®—
category_seq <- gene_anno_df$Category
rle_cat <- rle(as.character(category_seq))
gap_positions_genes <- cumsum(rle_cat$lengths)
if (length(gap_positions_genes) > 1) {
  gap_positions_genes <- gap_positions_genes[-length(gap_positions_genes)]
} else {
  gap_positions_genes <- NULL
}

cat("Gap positions (rows):", gap_positions_genes, "\n\n")

# ã‚«ãƒ†ã‚´ãƒªã‚«ãƒ©ãƒ¼
category_colors <- c(
  "T_cell_Exhaustion"       = "#984EA3",
  "Inflammatory_Chemokine"  = "#E41A1C",
  "Myeloid_Activation"      = "#FF7F00",
  "T_cell_Activation"       = "#A65628",
  "Tubular_Injury"          = "#FFD92F",
  "Endothelial_Activation"  = "#66C2A5",
  "Stromal_Activation"      = "#8DA0CB",
  "B_Plasma_Activation"     = "#FC8D62",
  "Tissue_Damage"           = "#E78AC3"
)

# annotation_colors ã« Category ã‚’è¿½åŠ 
ann_colors_full <- list(
  Cell_Type = celltype_cols,
  Condition = cond_cols,
  Category  = category_colors
)

# ã‚½ãƒ¼ãƒˆæ¸ˆã¿ matrix ã‚’ä½œæˆ
pseudobulk_scaled_sorted <- pseudobulk_scaled[gene_order, ]

# Heatmap
heat_cols <- colorRampPalette(c("navy", "white", "firebrick3"))(100)

p2c_heatmap <- pheatmap::pheatmap(
  mat               = pseudobulk_scaled_sorted,
  color             = heat_cols,
  breaks            = seq(-2.5, 2.5, length.out = 101),
  cluster_rows      = FALSE,
  cluster_cols      = FALSE,
  annotation_col    = anno_col_for_pheatmap,
  annotation_row    = gene_anno_df,
  annotation_colors = ann_colors_full,
  gaps_col          = gap_positions,
  gaps_row          = gap_positions_genes,
  show_colnames     = TRUE,
  show_rownames     = TRUE,
  fontsize_row      = 9,
  fontsize_col      = 7,
  angle_col         = 45,
  main              = "ICI-AKI signature genes by cell lineage (scFFPE)",
  border_color      = NA,
  cellwidth         = 15,
  cellheight        = 12,
  silent            = TRUE,
  labels_col        = labels_col
)

## ---- Save outputs ----
png(
  filename = file.path(output_dir,
                       "Figure2/Figure2C_scFFPE_signature_by_celltype_grouped.png"),
  width = 12, height = 8, units = "in", res = 300
)
print(p2c_heatmap)
dev.off()

pdf(
  file = file.path(output_dir,
                   "Figure2/Figure2C_scFFPE_signature_by_celltype_grouped.pdf"),
  width = 12, height = 8
)
print(p2c_heatmap)
dev.off()

write.csv(
  pseudobulk_matrix,
  file.path(output_dir, "Statistics/Figure2C_pseudobulk_expression_celltype_grouped.csv")
)

cat("âœ… Figure 2C (cell type-grouped) created and saved.\n\n")
}
}

## ---- figure2c_signature_heatmap_logFC_only, fig.width=12, fig.height=8 ----
cat("=== Creating Figure 2C (log2FC-ordered, no Category) ===\n")

# 1. ãƒ’ãƒ¼ãƒˆãƒžãƒƒãƒ—ã«ä½¿ã†éºä¼å­é›†åˆ
#   ï¼ˆæ—¢å­˜ã® Figure2C ã¨åŒã˜éºä¼å­ã«ã—ãŸã„å ´åˆï¼‰
genes_for_heatmap <- rownames(pseudobulk_scaled)  # ï¼å…ˆã»ã©ã® heatmap ã«ä½¿ã£ãŸéºä¼å­

# 2. ici_specific_genes ã® mean_log2FC ã§ä¸¦ã¹æ›¿ãˆ
gene_order_logfc <- ici_specific_genes %>%
  dplyr::filter(gene %in% genes_for_heatmap) %>%
  dplyr::arrange(dplyr::desc(mean_log2FC)) %>%
  dplyr::pull(gene)

# å¿µã®ãŸã‚ãƒ€ãƒ–ã‚Šãƒ»é †åºã‚’æ•´ãˆã‚‹
gene_order_logfc <- intersect(gene_order_logfc, genes_for_heatmap)

cat("Genes used (log2FC-ordered):\n")
print(gene_order_logfc)
cat("\n")

# 3. è¡Œé †ã‚’ log2FC é †ã«ä¸¦ã¹æ›¿ãˆãŸè¡Œåˆ—ã‚’ä½œæˆ
pseudobulk_scaled_logfc <- pseudobulk_scaled[gene_order_logfc, , drop = FALSE]

# 4. annotation_col ã¯æ—¢å­˜ã®ã‚‚ã®ã‚’å†åˆ©ç”¨ï¼ˆCell_Type / Condition ã®ã¿ï¼‰
anno_colors_logfc <- list(
  Cell_Type = celltype_cols,
  Condition = cond_cols
)

# 5. Heatmap ä½œæˆï¼ˆCategory ã® annotation_row ã¯ä»˜ã‘ãªã„ï¼‰
heat_cols <- colorRampPalette(c("navy", "white", "firebrick3"))(100)

p2c_heatmap_logfc <- pheatmap::pheatmap(
  mat               = pseudobulk_scaled_logfc,
  color             = heat_cols,
  breaks            = seq(-2.5, 2.5, length.out = 101),
  cluster_rows      = FALSE,
  cluster_cols      = FALSE,
  annotation_col    = anno_col_for_pheatmap,
  annotation_colors = anno_colors_logfc,
  gaps_col          = gap_positions,
  show_colnames     = TRUE,
  show_rownames     = TRUE,
  fontsize_row      = 9,
  fontsize_col      = 7,
  angle_col         = 45,
  main              = "ICI-AKI signature genes by cell lineage (log2FC-ordered, scFFPE)",
  border_color      = NA,
  cellwidth         = 15,
  cellheight        = 12,
  silent            = TRUE,
  labels_col        = labels_col
)

## 6. ä¿å­˜
png(
  filename = file.path(
    output_dir,
    "Figure2/Figure2C_scFFPE_signature_logFC_only.png"
  ),
  width = 12, height = 8, units = "in", res = 300
)
print(p2c_heatmap_logfc)
dev.off()

pdf(
  file = file.path(
    output_dir,
    "Figure2/Figure2C_scFFPE_signature_logFC_only.pdf"
  ),
  width = 12, height = 8
)
print(p2c_heatmap_logfc)
dev.off()

write.csv(
  pseudobulk_scaled_logfc,
  file.path(output_dir, "Statistics/Figure2C_pseudobulk_expression_logFC_only.csv")
)

cat("âœ… Figure 2C (log2FC-ordered, no Category) created and saved.\n\n")

```

## è¿½åŠ è§£æž: IFI27ã¨FOSã®ç™ºç¾ãƒ‘ã‚¿ãƒ¼ãƒ³

```{r ifi27_fos_visualization, fig.width=14, fig.height=10}
cat("=== Visualizing IFI27 and FOS Expression ===\n")

# å¯¾è±¡éºä¼å­
target_genes <- c("IFI27", "FOS")

# éºä¼å­ãŒãƒ‡ãƒ¼ã‚¿ã«å­˜åœ¨ã™ã‚‹ã‹ç¢ºèª
genes_available <- target_genes[target_genes %in% rownames(scffpe)]
genes_missing <- target_genes[!target_genes %in% rownames(scffpe)]

if (length(genes_missing) > 0) {
  cat("âš ï¸ Following genes not found in data:", paste(genes_missing, collapse = ", "), "\n")
}

if (length(genes_available) == 0) {
  cat("âŒ No target genes found in the dataset\n")
} else {
  cat("âœ… Found genes:", paste(genes_available, collapse = ", "), "\n\n")
  
  ## ---- 1. Violin Plot by Condition ----
  cat("Creating violin plots by condition...\n")
  
  vln_plots <- list()
  for (gene in genes_available) {
    p <- VlnPlot(
      scffpe,
      features = gene,
      group.by = "group",
      pt.size = 0,
      cols = condition_colors
    ) +
      labs(title = paste0(gene, " expression by condition"),
           x = "Condition",
           y = "Expression level") +
      theme_publication() +
      theme(legend.position = "none")
    
    vln_plots[[gene]] <- p
  }
  
  ## ---- 2. Violin Plot by Cell Type ----
  cat("Creating violin plots by cell type...\n")
  
  vln_celltype_plots <- list()
  for (gene in genes_available) {
    p <- VlnPlot(
      scffpe,
      features = gene,
      group.by = "major_celltype",
      pt.size = 0,
      cols = major_celltype_colors
    ) +
      labs(title = paste0(gene, " expression by cell type"),
           x = "Cell Type",
           y = "Expression level") +
      theme_publication() +
      theme(
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none"
      )
    
    vln_celltype_plots[[gene]] <- p
  }
  
  ## ---- 3. Feature Plot (UMAP) ----
  cat("Creating UMAP feature plots...\n")
  
  if ("umap" %in% names(scffpe@reductions)) {
    feature_plots <- FeaturePlot(
      scffpe,
      features = genes_available,
      reduction = "umap",
      ncol = 2,
      pt.size = 0.5,
      cols = c("lightgrey", "red")
    ) &
      theme_publication() &
      theme(
        legend.position = "right",
        plot.title = element_text(face = "bold", hjust = 0.5)
      )
  } else {
    cat("âš ï¸ UMAP reduction not found, skipping feature plots\n")
    feature_plots <- NULL
  }
  
  ## ---- 4. Dot Plot: Condition Ã— Cell Type ----
  cat("Creating dot plot...\n")
  
  # Condition Ã— Cell Type ã®çµ„ã¿åˆã‚ã›ã‚’ä½œæˆ
  scffpe$condition_celltype <- paste0(scffpe$group, "_", scffpe$major_celltype)
  
  dot_plot <- DotPlot(
    scffpe,
    features = genes_available,
    group.by = "condition_celltype",
    cols = c("lightgrey", "red"),
    dot.scale = 8
  ) +
    coord_flip() +
    labs(title = "Expression across Condition Ã— Cell Type") +
    theme_publication() +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
      axis.text.y = element_text(size = 10)
    )
  
  ## ---- 5. Heatmap: Pseudobulk by Condition ----
  cat("Creating pseudobulk heatmap...\n")
  
  # Pseudobulk by Condition
  pseudobulk_condition <- list()
  for (condition in unique(scffpe$group)) {
    cells <- rownames(scffpe@meta.data)[scffpe@meta.data$group == condition]
    if (length(cells) >= 10) {
      expr_data <- GetAssayData(scffpe, slot = "data", assay = "RNA")[genes_available, cells, drop = FALSE]
      pseudobulk_condition[[condition]] <- rowMeans(expr_data)
    }
  }
  
  if (length(pseudobulk_condition) > 0) {
    pb_matrix <- do.call(cbind, pseudobulk_condition)
    
    # Z-score normalization
    pb_scaled <- t(scale(t(pb_matrix)))
    pb_scaled[is.na(pb_scaled)] <- 0
    
    # Heatmap
    condition_anno <- data.frame(
      Condition = colnames(pb_scaled),
      row.names = colnames(pb_scaled)
    )
    
    anno_colors_cond <- list(
      Condition = condition_colors[colnames(pb_scaled)]
    )
    
    heatmap_condition <- pheatmap::pheatmap(
      pb_scaled,
      color = colorRampPalette(c("navy", "white", "firebrick3"))(100),
      breaks = seq(-2, 2, length.out = 101),
      cluster_rows = FALSE,
      cluster_cols = FALSE,
      annotation_col = condition_anno,
      annotation_colors = anno_colors_cond,
      show_colnames = TRUE,
      show_rownames = TRUE,
      fontsize_row = 12,
      fontsize_col = 10,
      main = "IFI27 and FOS: Pseudobulk by Condition",
      border_color = "black",
      cellwidth = 30,
      cellheight = 20
    )
  }
  
  ## ---- 6. Statistical Comparison ----
  cat("Performing statistical tests...\n")
  
  stat_results <- list()
  for (gene in genes_available) {
    # Extract expression data
    expr <- GetAssayData(scffpe, slot = "data", assay = "RNA")[gene, ]
    meta <- scffpe@meta.data
    meta$expression <- expr
    
    # Kruskal-Wallis test (non-parametric)
    kw_test <- kruskal.test(expression ~ group, data = meta)
    
    # Pairwise Wilcoxon test
    if (kw_test$p.value < 0.05) {
      pairwise <- pairwise.wilcox.test(
        meta$expression,
        meta$group,
        p.adjust.method = "BH"
      )
      
      stat_results[[gene]] <- list(
        kruskal_wallis = kw_test,
        pairwise = pairwise
      )
    }
  }
  
  ## ---- Save Outputs ----
  cat("\nSaving figures...\n")
  
  # 1. Violin plots by condition
  if (length(vln_plots) > 0) {
    combined_vln <- wrap_plots(vln_plots, ncol = 2)
    ggsave(
      filename = file.path(output_dir, "Additional/IFI27_FOS_violin_by_condition.png"),
      plot = combined_vln,
      width = 12, height = 6, dpi = 300
    )
    ggsave(
      filename = file.path(output_dir, "Additional/IFI27_FOS_violin_by_condition.pdf"),
      plot = combined_vln,
      width = 12, height = 6
    )
  }
  
  # 2. Violin plots by cell type
  if (length(vln_celltype_plots) > 0) {
    combined_vln_ct <- wrap_plots(vln_celltype_plots, ncol = 1)
    ggsave(
      filename = file.path(output_dir, "Additional/IFI27_FOS_violin_by_celltype.png"),
      plot = combined_vln_ct,
      width = 14, height = 10, dpi = 300
    )
    ggsave(
      filename = file.path(output_dir, "Additional/IFI27_FOS_violin_by_celltype.pdf"),
      plot = combined_vln_ct,
      width = 14, height = 10
    )
  }
  
  # 3. Feature plots
  if (!is.null(feature_plots)) {
    ggsave(
      filename = file.path(output_dir, "Additional/IFI27_FOS_UMAP.png"),
      plot = feature_plots,
      width = 12, height = 6, dpi = 300
    )
    ggsave(
      filename = file.path(output_dir, "Additional/IFI27_FOS_UMAP.pdf"),
      plot = feature_plots,
      width = 12, height = 6
    )
  }
  
  # 4. Dot plot
  ggsave(
    filename = file.path(output_dir, "Additional/IFI27_FOS_dotplot.png"),
    plot = dot_plot,
    width = 14, height = 6, dpi = 300
  )
  ggsave(
    filename = file.path(output_dir, "Additional/IFI27_FOS_dotplot.pdf"),
    plot = dot_plot,
    width = 14, height = 6
  )
  
  # 5. Pseudobulk heatmap
  if (exists("heatmap_condition")) {
    png(
      filename = file.path(output_dir, "Additional/IFI27_FOS_pseudobulk_heatmap.png"),
      width = 8, height = 6, units = "in", res = 300
    )
    print(heatmap_condition)
    dev.off()
    
    pdf(
      file = file.path(output_dir, "Additional/IFI27_FOS_pseudobulk_heatmap.pdf"),
      width = 8, height = 6
    )
    print(heatmap_condition)
    dev.off()
  }
  
  # 6. Save statistical results
  if (length(stat_results) > 0) {
    sink(file.path(output_dir, "Additional/IFI27_FOS_statistics.txt"))
    for (gene in names(stat_results)) {
      cat("\n==========================================\n")
      cat("Gene:", gene, "\n")
      cat("==========================================\n\n")
      
      cat("Kruskal-Wallis Test:\n")
      print(stat_results[[gene]]$kruskal_wallis)
      
      if (!is.null(stat_results[[gene]]$pairwise)) {
        cat("\nPairwise Wilcoxon Test (BH-adjusted):\n")
        print(stat_results[[gene]]$pairwise)
      }
      cat("\n")
    }
    sink()
  }
  
  ## ---- Summary Table ----
  cat("\nCreating summary table...\n")
  
  summary_data <- data.frame(
    Gene = character(),
    Condition = character(),
    Mean_Expression = numeric(),
    Median_Expression = numeric(),
    Pct_Expressing = numeric(),
    stringsAsFactors = FALSE
  )
  
  for (gene in genes_available) {
    expr <- GetAssayData(scffpe, slot = "data", assay = "RNA")[gene, ]
    
    for (condition in unique(scffpe$group)) {
      cells_in_cond <- scffpe$group == condition
      expr_cond <- expr[cells_in_cond]
      
      summary_data <- rbind(summary_data, data.frame(
        Gene = gene,
        Condition = condition,
        Mean_Expression = mean(expr_cond),
        Median_Expression = median(expr_cond),
        Pct_Expressing = sum(expr_cond > 0) / length(expr_cond) * 100
      ))
    }
  }
  
  write.csv(
    summary_data,
    file.path(output_dir, "Additional/IFI27_FOS_summary.csv"),
    row.names = FALSE
  )
  
  cat("\nâœ… IFI27 and FOS analysis complete!\n")
  cat("\nOutput files saved in:", file.path(output_dir, "Additional"), "\n\n")
  
  # Display summary table
  cat("Summary of expression:\n")
  print(summary_data)
}
```

## ã‚»ãƒƒã‚·ãƒ§ãƒ³æƒ…å ±

```{r session_info}
cat("=== SESSION INFORMATION ===\n")
sessionInfo()
```

---
